---
title: "Figure 07 - Panels ACDEF + FigS7"
output: 
  rmdformats::robobook:
    lightbox: true
    fig_caption: true
---

<style type="text/css">
  body{
  font-size: 10pt;
}
</style>

# 1) Setup

## - Defining work directory

In this section, we define the working directory for the R Markdown document.

```{r setup, include=TRUE}
# Get the path of the current script
# Then get the parent directory of the parent directory of the parent directory
local_wd_folder <- dirname(dirname(dirname(rstudioapi::getSourceEditorContext()$path)))

# Set the root directory for knitr to the local working directory
knitr::opts_knit$set(root.dir = local_wd_folder)
```

## - Defining input data and output directories

Here, we define the directories for input data and output files.

```{r}
# Get the directory of the current script
script_folder <- dirname(rstudioapi::getSourceEditorContext()$path)

# Define the data folder and output folder
data_folder <- './Data'
#results_folder <- './Results'
figures_folder <- './Figures'

dir.create(paste(data_folder, '/RandomForest/Results/Clustering/K5', sep = ''), showWarnings = FALSE)
```

## - Setting seed

Setting a seed ensures that any random processes are reproducible.

```{r}
# Set a seed for reproducibility
set.seed(123)
```

## - Packages installation (optional)

In this section, we ensure that all necessary packages are installed.

```{r}
# Ensure BiocManager is available for installation of Bioconductor packages
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
#BiocManager::install(version = "3.18")

# Define a list of required packages used in this script
packages_required <- c("ggplot2", "dplyr", "coin", "randomForest", "rstatix",
                       "Boruta", "rattle", "caret", "ROCR", "pROC", "dendextend")

# Identify any required packages that are not installed
packages_uninstalled <- packages_required[!(packages_required %in% installed.packages()[,"Package"])]

# Install any uninstalled packages
if(length(packages_uninstalled)) BiocManager::install(packages_uninstalled)
```

## - Loading packages

Here, we load the necessary packages for our analysis.

```{r message=FALSE, warning=FALSE}
library(ggplot2, quietly = TRUE)
library(randomForest, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(coin, quietly = TRUE)
library(rstatix, quietly = TRUE)
library(Boruta, quietly = TRUE)
library(rattle, quietly = TRUE)
library(caret, quietly = TRUE)
library(ROCR, quietly = TRUE)
library(pROC, quietly = TRUE)
library(dendextend, quietly = TRUE)
```

## - Loading palettes

In this section, we load additional color palettes and define some custom ones.

```{r message=FALSE, warning=FALSE}
# Load additional colour palette packages
library(unikn, quietly = TRUE)
library(RColorBrewer, quietly = TRUE)
library(yarrr, quietly = TRUE)
library(scales, quietly = TRUE)
library(ggsci, quietly = TRUE)

# Define a set of custom color palettes from the unikn package
mix_1 <- usecol(pal = c(Karpfenblau, "white", Peach), n = 15)
mix_2 <- usecol(pal = c(rev(pal_seeblau), "white", pal_pinky))
mix_3 <- usecol(pal = c(rev(pal_bordeaux), "white", pal_petrol), n = 15)

# Display the custom palettes
seecol(list(mix_1, mix_2, mix_3), col_brd = "white", lwd_brd = 4, title = "Comparing palettes mixed from unikn colors", pal_names = c("mix_1", "mix_2", "mix_3"))

# Define a second set of custom palettes from the RColorBrewer and yarrr packages
brew_mix <- usecol(c(rev(brewer.pal(n = 4, name = "Reds")), "white", brewer.pal(n = 4, name = "Blues")), n = 13)
brew_ext <- usecol(brewer.pal(n = 11, name = "Spectral"), n = 12)
yarrr_mix <- usecol(c(piratepal("nemo"), piratepal("bugs")))
yarrr_mod <- usecol(c(piratepal("ipod")), n = 9)

# Display the second set of custom palettes
seecol(pal = list(brew_mix, brew_ext, yarrr_mix, yarrr_mod), col_brd = "white", lwd_brd = 2, title = "Using usecol() and seecol() to mix and modify palettes", pal_names = c("brew_mix", "brew_ext", "yarrr_mix", "yarrr_mod"))

# Define additional custom palettes from the scales package
natjournals_palette <- pal_npg("nrc")(10)
```

## - Log Session Info

Finally, we log the session information for reproducibility.

```{r message=TRUE, warning=TRUE}
# Write the session information to a text file
writeLines(capture.output(sessionInfo()), file.path(script_folder, 'Fig_07_ACDEF_Fig_S7_SessionInfo.txt'))

# Print the session information
sessionInfo()
```

# 2) Loading input files

## - Loading K2 RF clusters

```{r}
RF_cluster_metadata <- read.csv(paste(data_folder, '/RandomForest/Results/Clustering/K2/RF_K2_Clustered_Metadata.csv', 
                                    sep = ''), row.names = 'Patient.ID')
```

# 3) Vizome BeatAML drug responses heatmap - K2 : Fig S7

```{r}
BeatAML_drug_response <- as.matrix(RF_cluster_metadata[,91:177])

colnames(BeatAML_drug_response) <- c(
                                "A-674563",
                                "ABT-737",
                                "Afatinib (BIBW-2992)",
                                "AT7519",
                                "Axitinib (AG-013736)",
                                "AZD1152-HQPA (AZD2811)",
                                "AZD1480",
                                "Bay 11-7085",
                                "BEZ235",
                                "Birinapant",
                                "Bortezomib (Velcade)",
                                "Bosutinib (SKI-606)",
                                "Cabozantinib",
                                "CI-1040 (PD184352)",
                                "Crenolanib",
                                "Crizotinib (PF-2341066)",
                                "CYT387",
                                "Dasatinib",
                                "Doramapimod (BIRB 796)",
                                "Dovitinib (CHIR-258)",
                                "Elesclomol",
                                "Entospletinib (GS-9973)",
                                "Entrectinib",
                                "Flavopiridol",
                                "Foretinib (XL880)",
                                "GDC-0941",
                                "Gilteritinib",
                                "GSK-1838705A",
                                "GW-2580",
                                "H-89",
                                "Ibrutinib (PCI-32765)",
                                "Idelalisib",
                                "Indisulam",
                                "INK-128",
                                "JAK Inhibitor I",
                                "JNJ-28312141",
                                "JNJ-38877605",
                                "JQ1",
                                "KI20227",
                                "KU-55933",
                                "KW-2449",
                                "Lenalidomide",
                                "Lenvatinib",
                                "Linifanib (ABT-869)",
                                "Masitinib (AB-1010)",
                                "MGCD-265",
                                "Midostaurin",
                                "MK-2206",
                                "MLN120B",
                                "Neratinib (HKI-272)",
                                "NF-kB Activation Inhibitor",
                                "Nutlin 3a",
                                "NVP-ADW742",
                                "NVP-TAE684",
                                "OTX-015",
                                "Panobinostat",
                                "PD173955",
                                "Pelitinib (EKB-569)",
                                "Perhexiline maleate",
                                "PH-797804",
                                "PHA-665752",
                                "PI-103",
                                "Ponatinib (AP24534)",
                                "PP242",
                                "PRT062607",
                                "Quizartinib (AC220)",
                                "RAF265 (CHIR-265)",
                                "Ralimetinib (LY2228820)",
                                "Ranolazine",
                                "Rapamycin",
                                "Regorafenib (BAY 73-4506)",
                                "S31-201",
                                "Selinexor",
                                "Selumetinib (AZD6244)",
                                "SNS-032 (BMS-387032)",
                                "Sorafenib",
                                "Sunitinib",
                                "Tivozanib (AV-951)",
                                "Tofacitinib (CP-690550)",
                                "Tozasertib (VX-680)",
                                "Trametinib (GSK1120212)",
                                "Vandetanib (ZD6474)",
                                "Vargetef",
                                "Vatalanib (PTK787)",
                                "Volasertib (BI-6727)",
                                "VX-745",
                                "Tanespimycin (17-AAG)")

Identity <- data.frame(row.names = rownames(RF_cluster_metadata))

Identity$RF_Cluster <- factor(RF_cluster_metadata$RF_Cluster, 
                                         levels=c('C1',
                                                  'C2'))

colors_annotation <-  list(RF_Cluster = c('C1' = '#793F0B',
                                          'C2' = '#0B544B'))

row_anno <- rowAnnotation(df=Identity, col = colors_annotation, 
                        gp = gpar(lwd =0.25, col = 'black'), simple_anno_size = unit(5, "mm"),
                        show_annotation_name = F, show_legend = FALSE,
                        annotation_legend_param = list(RF_Cluster = list(direction = "vertical")))

breaks_val <- seq(from = 0, to = 1, length.out = 14)
col_heatmap_guide <- colorRamp2(breaks = breaks_val,colors = rev(c("#67000D",brew_mix)))

HMap_BeatAMLdrugs <- Heatmap(BeatAML_drug_response, 
        name = 'ssGSEA Score',
        row_split = RF_cluster_metadata$RF_Cluster, 
        left_annotation  = row_anno,
        col = col_heatmap_guide, 
        rect_gp = gpar(col = "black", lwd = 0.05),
        cluster_row_slices = F,
        column_title_gp = gpar(fontsize=14, fontface='bold'),
        row_names_gp = gpar(fontsize=8), 
        cluster_rows = T,
        column_names_gp = gpar(fontsize=8), 
        show_heatmap_legend = T,
        heatmap_legend_param = list(legend_direction = "horizontal",
                                    legend_width = unit(5, "cm")),
        row_names_side = 'right')

draw(HMap_BeatAMLdrugs, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")

png(filename = paste(figures_folder, '/Supplemental/Fig_S7.png', sep = ''), units = "cm", width = 30, height = 20, res = 300)

draw(HMap_BeatAMLdrugs, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")

invisible(dev.off())
```

# 4) RandomForest Subclustering

## - Loading RandomForest model

```{r}
model <- readRDS(file = paste(data_folder, '/RandomForest/Results/Model/RF_Classification_Model.RDS', sep = ''))
```

## - Analysis of RF model

### - Extracting proximity matrix for distance computation and MDS similarity evaluation

```{r}
distance.matrix <- dist(1-model$proximity)

mds.stuff <- cmdscale(distance.matrix, eig = TRUE, x.ret = TRUE)
mds.var.per <- round(mds.stuff$eig/sum(mds.stuff$eig)*100, 1)
mds.values <- mds.stuff$points
mds.data <- data.frame(Sample=rownames(mds.values),
                       X= mds.values[,1],
                       Y= mds.values[,2])

ggplot(data = mds.data, aes(x = X, y = Y, label=Sample)) + 
  geom_text(aes(color=Sample)) +
  theme_bw() +
  xlab(paste("MDS1 - ", mds.var.per[1], "%", sep="")) +
  ylab(paste("MDS2 - ", mds.var.per[2], "%", sep="")) +
  ggtitle(label = "MDS plot", subtitle = "(1 - Random Forest Proximities)") + theme(legend.position = 'bottom')

```

### - Clustering K = 5

#### - Hierarchical clustering on distance matrix : Panel A

```{r}
dendro <- stats::hclust(distance.matrix, method = "ward.D2") %>% as.dendrogram %>% ladderize  %>% color_branches(k = 5, col = c('#C67942', '#E0B090', '#87551B', '#7CB7B0', '#539690')) %>% color_labels(k = 5, col = c('#C67942', '#E0B090', '#87551B', '#7CB7B0', '#539690'))

plot(dendro, axes=F)
par(lwd=3)

#


Hierchical_clusters_K5 <- sort(cutree(dendro, k = 5))

Hierarchical_clustering_K5 <- data.frame(row.names = rownames(RF.metadata))
Hierarchical_clustering_K5$Patient.ID <- rownames(Hierarchical_clustering_K5)

for (ID in Hierarchical_clustering_K5$Patient.ID) {
  
Hierarchical_clustering_K5[ID,'Cluster'] <- Hierchical_clusters_K5[[ID]]

}

Hierarchical_clustering_K5[Hierarchical_clustering_K5$Cluster == Hierarchical_clustering_K5['671', 'Cluster'],'Cluster_annotated'] <- 'C1A: CD34- HSC/MPP primed'
Hierarchical_clustering_K5[Hierarchical_clustering_K5$Cluster == Hierarchical_clustering_K5['543', 'Cluster'],'Cluster_annotated'] <- 'C1B: CD34+ HSC/MPP primed'
Hierarchical_clustering_K5[Hierarchical_clustering_K5$Cluster == Hierarchical_clustering_K5['544', 'Cluster'],'Cluster_annotated'] <- 'C1C: Myeloid Prog. primed'
Hierarchical_clustering_K5[Hierarchical_clustering_K5$Cluster == Hierarchical_clustering_K5['303', 'Cluster'],'Cluster_annotated'] <- 'C2A: Promonocyte primed'
Hierarchical_clustering_K5[Hierarchical_clustering_K5$Cluster == Hierarchical_clustering_K5['55', 'Cluster'],'Cluster_annotated'] <- 'C2B: Monocyte primed'

mds.data.Hierarchical_K5_clustered <- mds.data

mds.data.Hierarchical_K5_clustered$Cluster <- Hierarchical_clustering_K5$Cluster_annotated

cols <- c('C1B: CD34+ HSC/MPP primed' = '#C67942', 'C1C: Myeloid Prog. primed' = '#E0B090', 'C1A: CD34- HSC/MPP primed' = '#87551B', 
          'C2A: Promonocyte primed' = '#7CB7B0', 'C2B: Monocyte primed' = '#539690')

plot.Hierarchical_clusters_K5 <- ggplot(data = mds.data.Hierarchical_K5_clustered, aes(x = X, y = Y, label=Sample, color=Cluster)) + 
  geom_point(aes(color=Cluster, fill=Cluster), shape = 21, alpha = 1, size = 2.5) +
  theme_linedraw(base_line_size = 0.2) +
  stat_ellipse(data = mds.data.Hierarchical_K5_clustered, geom = "polygon",
               aes(fill = Cluster), alpha = 0.15, linetype = 2) +
  xlab("MDS1") +
  ylab("MDS2") +
  ggtitle(label = "") +
  scale_colour_manual(
    values = cols,
    aesthetics = c("colour", "fill")
  ) + theme(legend.position = 'none',
            axis.title.x = element_text(family='Helvetica', size=12, face = 'bold', vjust = 0),
            axis.title.y = element_text(family='Helvetica', size=12, face = 'bold', vjust = +3))

print(plot.Hierarchical_clusters_K5)

png(filename = paste(figures_folder, '/Fig_07/Panel_A.png', sep = ''), units = "cm", width = 14, height = 14, res = 300)

print(plot.Hierarchical_clusters_K5)

invisible(dev.off())

#

plot.Hierarchical_clusters_K5_labelled <- plot.Hierarchical_clusters_K5 + geom_label_repel(aes(label = Sample, color=Cluster),
                  box.padding   = 0.65, 
                  point.padding = 0.1, size = 2,
                  segment.color = 'grey70', max.overlaps = 15)

print(plot.Hierarchical_clusters_K5_labelled) 

#

RF.metadata.hierarchical_clustering_K5 <- RF.metadata
RF.metadata.hierarchical_clustering_K5[Hierarchical_clustering_K5$Patient.ID,'RF_Cluster'] <- Hierarchical_clustering_K5$Cluster_annotated

RF.metadata.hierarchical_clustering_K5 <- RF.metadata.hierarchical_clustering_K5 %>% add_column(.before = 'Gender', 'Patient.ID' = rownames(RF.metadata.hierarchical_clustering_K5))

RF.metadata.hierarchical_clustering_K5[rownames(RF.metadata.hierarchical_clustering_K5 %>% filter(RF_Cluster == 'C1A: CD34- HSC/MPP primed')),'RF_Cluster_short'] <- 'C1A'
RF.metadata.hierarchical_clustering_K5[rownames(RF.metadata.hierarchical_clustering_K5 %>% filter(RF_Cluster == 'C1B: CD34+ HSC/MPP primed')),'RF_Cluster_short'] <- 'C1B'
RF.metadata.hierarchical_clustering_K5[rownames(RF.metadata.hierarchical_clustering_K5 %>% filter(RF_Cluster == 'C1C: Myeloid Prog. primed')),'RF_Cluster_short'] <- 'C1C'
RF.metadata.hierarchical_clustering_K5[rownames(RF.metadata.hierarchical_clustering_K5 %>% filter(RF_Cluster == 'C2A: Promonocyte primed')),'RF_Cluster_short'] <- 'C2A'
RF.metadata.hierarchical_clustering_K5[rownames(RF.metadata.hierarchical_clustering_K5 %>% filter(RF_Cluster == 'C2B: Monocyte primed')),'RF_Cluster_short'] <- 'C2B'

RF.metadata.hierarchical_clustering_K5$RF_Cluster_short <- factor(RF.metadata.hierarchical_clustering_K5$RF_Cluster_short)

write_csv(x = RF.metadata.hierarchical_clustering_K5, 
            file = paste(data_folder, '/RandomForest/Results/Clustering/K5/RF_K5_Clustered_Metadata.csv', sep = ''))

RF.metadata.hierarchical_clustering_K5$Patient.ID <- NULL
```

#### - Wilcox t-test unpaired - median comparisons : Panel B

Features were retrospectively compared in the identified groupings. Selected features were illustrated via GraphPad (see under ***GraphPad/Fig_07/Panels_B.pzfx***).

```{r}
RF.metadata.hierarchical_clustering_K5$RF_Cluster <- NULL # We remove this column for the following analysis

RF.metadata.hierarchical_clustering_K5_numerised <- RF.metadata.hierarchical_clustering_K5[1:length(colnames(RF.metadata.hierarchical_clustering_K5))]

for (cluster1 in unique(RF.metadata.hierarchical_clustering_K5_numerised$RF_Cluster_short)) {

RF.metadata.hierarchical_clustering_K5_numerised <- RF.metadata.hierarchical_clustering_K5[1:length(colnames(RF.metadata.hierarchical_clustering_K5))]

  for (cluster2 in unique(RF.metadata.hierarchical_clustering_K5_numerised$RF_Cluster_short)[unique(RF.metadata.hierarchical_clustering_K5_numerised$RF_Cluster_short)!= cluster1]) {

RF.metadata.hierarchical_clustering_K5_numerised <- RF.metadata.hierarchical_clustering_K5[1:length(colnames(RF.metadata.hierarchical_clustering_K5))]

  RF.metadata.hierarchical_clustering_K5_numerised <- RF.metadata.hierarchical_clustering_K5_numerised %>% filter(RF_Cluster_short == cluster1 | RF_Cluster_short == cluster2)
  
  variable_to_compare <- c(colnames(RF.metadata.hierarchical_clustering_K5_numerised)[1:length(colnames(RF.metadata.hierarchical_clustering_K5))])
  
  for (variable in colnames(RF.metadata.hierarchical_clustering_K5_numerised)[1:length(colnames(RF.metadata.hierarchical_clustering_K5))-1]) {
  
  RF.metadata.hierarchical_clustering_K5_numerised[,variable] <- as.numeric(RF.metadata.hierarchical_clustering_K5_numerised[,variable])
    
  }
  
  Group_A <- RF.metadata.hierarchical_clustering_K5_numerised %>% dplyr::filter(RF_Cluster_short == cluster1)
  Group_B <- RF.metadata.hierarchical_clustering_K5_numerised %>% dplyr::filter(RF_Cluster_short == cluster2)
  
  wilcox_t_test_results <- data.frame(row.names = colnames(RF.metadata.hierarchical_clustering_K5_numerised)[1:length(colnames(RF.metadata.hierarchical_clustering_K5))-1])
  
  for (variable in colnames(RF.metadata.hierarchical_clustering_K5_numerised)[1:length(colnames(RF.metadata.hierarchical_clustering_K5))-1]) {
    if (((length(unique(RF.metadata.hierarchical_clustering_K5_numerised[,variable]))) == 1) == TRUE) {
    wilcox_t_test_results[variable,'pval'] <- 1
    wilcox_t_test_results[variable, paste('Median', cluster1, sep = ' ')] <- median(Group_A[,variable])
    wilcox_t_test_results[variable, paste('Median', cluster2, sep = ' ')] <- median(Group_B[,variable])
    wilcox_t_test_results[variable, paste('Min', cluster1, sep = ' ')] <- min(Group_A[,variable])
    wilcox_t_test_results[variable, paste('Min', cluster2, sep = ' ')] <- min(Group_B[,variable])
    wilcox_t_test_results[variable, paste('Max', cluster1, sep = ' ')] <- max(Group_A[,variable])
    wilcox_t_test_results[variable, paste('Max', cluster2, sep = ' ')] <- max(Group_B[,variable])
    wilcox_t_test_results[variable, paste('Std', cluster1, sep = ' ')] <- sd(Group_A[,variable])
    wilcox_t_test_results[variable, paste('Std', cluster2, sep = ' ')] <- sd(Group_B[,variable])
    } else {
    p <- coin::wilcox_test(formula = get(variable) ~ as.factor(RF_Cluster_short), data=RF.metadata.hierarchical_clustering_K5_numerised, distribution = "exact")
    wilcox_t_test_results[variable,'pval'] <- coin::pvalue(p)
    wilcox_t_test_results[variable, paste('Median', cluster1, sep = ' ')] <- median(Group_A[,variable])
    wilcox_t_test_results[variable, paste('Median', cluster2, sep = ' ')] <- median(Group_B[,variable])
    wilcox_t_test_results[variable, paste('Min', cluster1, sep = ' ')] <- min(Group_A[,variable])
    wilcox_t_test_results[variable, paste('Min', cluster2, sep = ' ')] <- min(Group_B[,variable])
    wilcox_t_test_results[variable, paste('Max', cluster1, sep = ' ')] <- max(Group_A[,variable])
    wilcox_t_test_results[variable, paste('Max', cluster2, sep = ' ')] <- max(Group_B[,variable])
    wilcox_t_test_results[variable, paste('Std', cluster1, sep = ' ')] <- sd(Group_A[,variable])
    wilcox_t_test_results[variable, paste('Std', cluster2, sep = ' ')] <- sd(Group_B[,variable])
    }
  }
  
  wilcox_t_test_results <- wilcox_t_test_results %>% arrange(pval)
  
  write.csv(wilcox_t_test_results, 
            file = paste(data_folder, '/RandomForest/Results/Clustering/K5/RF_K5_Stats_Wilcox_t_test__',
                         cluster1, '_vs_', cluster2, '.csv', sep = ''),
            row.names = T, quote = F)
  }
}

RF.metadata.hierarchical_clustering_K5_numerised <- RF.metadata.hierarchical_clustering_K5[1:length(colnames(RF.metadata.hierarchical_clustering_K5))]
```

# 5) Distribution VanGalen AML signatures enrichment scores : Panel C

## - Loading K5 RF clusters

```{r}
RF_cluster_metadata <- read.csv(paste(data_folder, '/RandomForest/Results/Clustering/K5/RF_K5_Clustered_Metadata.csv', 
                                    sep = ''), row.names = 'Patient.ID')
```

```{r}
VanGalen_celltype_enrichment_scores <- as.matrix(RF_cluster_metadata[,79:86])

colnames(VanGalen_celltype_enrichment_scores) <- c('HSC-like',
                                     'HSC.Prog-like',
                                     'Progenitor-like',
                                     'GMP-like',
                                     'Promono-like',
                                     'Monocyte-like',
                                     'cDC-like',
                                     'Myeloid-like')

VanGalen_celltype_enrichment_scores <- t(VanGalen_celltype_enrichment_scores)

Identity <- data.frame(row.names = rownames(RF_cluster_metadata))

Identity$RF_Cluster <- factor(RF_cluster_metadata$RF_Cluster_short, 
                                         levels=c('C1A',
                                                  'C1C',
                                                  'C1B',
                                                  'C2A',
                                                  'C2B'))

colors_annotation <-  list(RF_Cluster = c('C1A' = '#87551B',
                                        'C1C' = '#E0B090', 
                                        'C1B' = '#C67942', 
                                        'C2A' = '#7CB7B0', 
                                        'C2B' = '#539690'))

col_anno <- HeatmapAnnotation(df=Identity, col = colors_annotation, 
                        gp = gpar(lwd =0.25, col = 'black'), simple_anno_size = unit(2, "mm"),
                        show_annotation_name = F, show_legend = FALSE,
                        annotation_legend_param = list(RF_Cluster = list(direction = "vertical")))

breaks_val <- seq(from = 0, to = 1, length.out = 14)
col_heatmap_guide <- colorRamp2(breaks = breaks_val,colors = rev(c("#67000D",brew_mix)))

HMap_vangalen <- Heatmap(VanGalen_celltype_enrichment_scores, 
        name = 'ssGSEA Score',
        column_split = RF_cluster_metadata$RF_Cluster_short, top_annotation = col_anno,
        col = col_heatmap_guide, rect_gp = gpar(col = "black", lwd = 0.05),
        column_title_gp = gpar(fontsize=11, fontface='bold'),
        row_names_gp = gpar(fontsize=10), cluster_column_slices  = F, cluster_rows = F,
        column_names_gp = gpar(fontsize=10), show_heatmap_legend = F,
        heatmap_legend_param = list(legend_direction = "horizontal",
                                    legend_height = unit(5, "cm")),
        row_names_side = 'right')

draw(HMap_vangalen, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")


png(filename = paste(figures_folder, '/Fig_07/Panel_C.png', sep = ''), units = "cm", width = 18.62, height = 7, res = 300)

draw(HMap_vangalen, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")

invisible(dev.off())
```

# 6) Vizome BeatAML drug responses heatmap - K5 : Panel D

```{r}
BeatAML_drug_response <- as.matrix(RF_cluster_metadata[,91:177])

colnames(BeatAML_drug_response) <- c(
                                "A-674563",
                                "ABT-737",
                                "Afatinib (BIBW-2992)",
                                "AT7519",
                                "Axitinib (AG-013736)",
                                "AZD1152-HQPA (AZD2811)",
                                "AZD1480",
                                "Bay 11-7085",
                                "BEZ235",
                                "Birinapant",
                                "Bortezomib (Velcade)",
                                "Bosutinib (SKI-606)",
                                "Cabozantinib",
                                "CI-1040 (PD184352)",
                                "Crenolanib",
                                "Crizotinib (PF-2341066)",
                                "CYT387",
                                "Dasatinib",
                                "Doramapimod (BIRB 796)",
                                "Dovitinib (CHIR-258)",
                                "Elesclomol",
                                "Entospletinib (GS-9973)",
                                "Entrectinib",
                                "Flavopiridol",
                                "Foretinib (XL880)",
                                "GDC-0941",
                                "Gilteritinib",
                                "GSK-1838705A",
                                "GW-2580",
                                "H-89",
                                "Ibrutinib (PCI-32765)",
                                "Idelalisib",
                                "Indisulam",
                                "INK-128",
                                "JAK Inhibitor I",
                                "JNJ-28312141",
                                "JNJ-38877605",
                                "JQ1",
                                "KI20227",
                                "KU-55933",
                                "KW-2449",
                                "Lenalidomide",
                                "Lenvatinib",
                                "Linifanib (ABT-869)",
                                "Masitinib (AB-1010)",
                                "MGCD-265",
                                "Midostaurin",
                                "MK-2206",
                                "MLN120B",
                                "Neratinib (HKI-272)",
                                "NF-kB Activation Inhibitor",
                                "Nutlin 3a",
                                "NVP-ADW742",
                                "NVP-TAE684",
                                "OTX-015",
                                "Panobinostat",
                                "PD173955",
                                "Pelitinib (EKB-569)",
                                "Perhexiline maleate",
                                "PH-797804",
                                "PHA-665752",
                                "PI-103",
                                "Ponatinib (AP24534)",
                                "PP242",
                                "PRT062607",
                                "Quizartinib (AC220)",
                                "RAF265 (CHIR-265)",
                                "Ralimetinib (LY2228820)",
                                "Ranolazine",
                                "Rapamycin",
                                "Regorafenib (BAY 73-4506)",
                                "S31-201",
                                "Selinexor",
                                "Selumetinib (AZD6244)",
                                "SNS-032 (BMS-387032)",
                                "Sorafenib",
                                "Sunitinib",
                                "Tivozanib (AV-951)",
                                "Tofacitinib (CP-690550)",
                                "Tozasertib (VX-680)",
                                "Trametinib (GSK1120212)",
                                "Vandetanib (ZD6474)",
                                "Vargetef",
                                "Vatalanib (PTK787)",
                                "Volasertib (BI-6727)",
                                "VX-745",
                                "Tanespimycin (17-AAG)")

Identity <- data.frame(row.names = rownames(RF_cluster_metadata))

Identity$RF_Cluster <- factor(RF_cluster_metadata$RF_Cluster_short, 
                                         levels=c('C1B',
                                                  'C1C',
                                                  'C1A',
                                                  'C2A',
                                                  'C2B'))

colors_annotation <-  list(RF_Cluster = c('C1A' = '#87551B',
                                        'C1C' = '#E0B090', 
                                        'C1B' = '#C67942', 
                                        'C2A' = '#7CB7B0', 
                                        'C2B' = '#539690'))

row_anno <- rowAnnotation(df=Identity, col = colors_annotation, 
                        gp = gpar(lwd =0.25, col = 'black'), simple_anno_size = unit(5, "mm"),
                        show_annotation_name = F, show_legend = FALSE,
                        annotation_legend_param = list(RF_Cluster = list(direction = "vertical")))

breaks_val <- seq(from = 0, to = 1, length.out = 14)
col_heatmap_guide <- colorRamp2(breaks = breaks_val,colors = rev(c("#67000D",brew_mix)))

dendro <- stats::hclust(dist(t(BeatAML_drug_response)), method = 'ward.D') %>% as.dendrogram %>% ladderize

plot(dendro, axes=F)
par(lwd=3)

Hierchical_clusters_drugs <- sort(cutree(dendro, k = 3))

drug_df <- data.frame(row.names = names(Hierchical_clusters_drugs))

for (drug in rownames(drug_df)) {

drug_df[drug,'Hierarchical_cluster'] <- Hierchical_clusters_drugs[[drug]]

}

drug_df$Hierarchical_cluster <- factor(drug_df$Hierarchical_cluster, levels = c('1','2','3'))

HMap_BeatAMLdrugs <- Heatmap(BeatAML_drug_response[,rownames(drug_df)],
        name = 'ssGSEA Score',
        row_split = RF_cluster_metadata$RF_Cluster_short,
        left_annotation  = row_anno,
        column_split = drug_df$Hierarchical_cluster,
        col = col_heatmap_guide,
        rect_gp = gpar(col = "black", lwd = 0.05),
        cluster_row_slices = F,
        cluster_column_slices = F,
        column_title_gp = gpar(fontsize=0, fontface='bold'),
        row_names_gp = gpar(fontsize=9),
        column_gap = unit(c(2.5,2.5), "mm"),
        cluster_rows = T,
        cluster_columns = T,
        column_names_gp = gpar(fontsize=9),
        show_heatmap_legend = F,
        heatmap_legend_param = list(legend_direction = "horizontal",
                                    legend_width = unit(5, "cm")),
        row_names_side = 'right')

draw(HMap_BeatAMLdrugs, heatmap_legend_side = "bottom",
     annotation_legend_side = "bottom")


png(filename = paste(figures_folder, '/Fig_07/Panel_D.png', sep = ''), units = "cm", width = 30, height = 22, res = 300)

draw(HMap_BeatAMLdrugs, heatmap_legend_side = "bottom",
     annotation_legend_side = "bottom")

invisible(dev.off())
```

# 7) Vizome BeatAML drug responses grouped by family heatmap - K5 : Panel E

```{r}
Drug_families <- read.csv(paste(data_folder, '/RNASeq/PubliclyAvailable/2022__Vizome__BeatAML/Metadata/Drug_families.txt', sep = ''),
                          sep = '\t')

Drug_families <- Drug_families %>% filter(inhibitor %in% colnames(BeatAML_drug_response))

Unique_families <- levels(factor(Drug_families$family))[grep(TRUE,table(Drug_families$family) > 1)]

Unique_families_df <- data.frame(row.names = Unique_families)

Drug_families_short_names <- read.csv(paste(data_folder, '/RNASeq/PubliclyAvailable/2022__Vizome__BeatAML/Metadata/Drug_families_short_names.txt',
                                            sep = ''),
                          sep = '\t')

for (unique_family in Unique_families) {
Unique_families_df[unique_family,'Short_name'] <- (Drug_families_short_names %>% filter(family == unique_family))$synonym
}

family_sensitivity_score <- data.frame(row.names = rownames(BeatAML_drug_response))

for (unique_family in Unique_families) {
  
family_sensitivity_score[,Unique_families_df[unique_family,'Short_name']] <- matrixStats::rowMedians(BeatAML_drug_response[,(Drug_families %>% filter(family == unique_family))$inhibitor])

}


Identity <- data.frame(row.names = rownames(RF_cluster_metadata))

Identity$RF_Cluster <- factor(RF_cluster_metadata$RF_Cluster_short, 
                                         levels=c('C1A',
                                                  'C1C',
                                                  'C1B',
                                                  'C2A',
                                                  'C2B'))

colors_annotation <-  list(RF_Cluster = c('C1A' = '#87551B',
                                        'C1C' = '#E0B090', 
                                        'C1B' = '#C67942', 
                                        'C2A' = '#7CB7B0', 
                                        'C2B' = '#539690'))

row_anno <- rowAnnotation(df=Identity, col = colors_annotation, 
                        gp = gpar(lwd =0.25, col = 'black'), simple_anno_size = unit(5, "mm"),
                        show_annotation_name = F, show_legend = FALSE,
                        annotation_legend_param = list(RF_Cluster = list(direction = "vertical")))

breaks_val <- seq(from = 0, to = 1, length.out = 14)
col_heatmap_guide <- colorRamp2(breaks = breaks_val,colors = rev(c("#67000D",brew_mix)))


dendro <- stats::hclust(dist(t(family_sensitivity_score)), method = 'ward.D') %>% as.dendrogram %>% ladderize

plot(dendro, axes=F)
par(lwd=3)

Hierchical_clusters_drugs <- sort(cutree(dendro, k = 3))

drug_family_df <- data.frame(row.names = names(Hierchical_clusters_drugs))

for (drug_family in rownames(drug_family_df)) {
  
drug_family_df[drug_family,'Hierarchical_cluster'] <- Hierchical_clusters_drugs[[drug_family]]
               
}

drug_family_df$Hierarchical_cluster <- factor(drug_family_df$Hierarchical_cluster, levels = c('3','1','2'))



HMap_BeatAMLdrugs <- Heatmap(family_sensitivity_score[,rownames(drug_family_df)], 
        name = 'ssGSEA Score',
        row_split = RF_cluster_metadata$RF_Cluster_short, 
        left_annotation  = row_anno,
        column_split = drug_family_df$Hierarchical_cluster,
        col = col_heatmap_guide, 
        rect_gp = gpar(col = "black", lwd = 0.05),
        cluster_row_slices = F,
        cluster_column_slices = F,
        column_title_gp = gpar(fontsize=0, fontface='bold'),
        row_names_gp = gpar(fontsize=9), 
        column_gap = unit(c(2.5,2.5), "mm"),
        cluster_rows = T,
        cluster_columns = T,
        column_names_gp = gpar(fontsize=9), 
        show_heatmap_legend = F,
        heatmap_legend_param = list(legend_direction = "horizontal",
                                    legend_width = unit(5, "cm")),
        row_names_side = 'right')

draw(HMap_BeatAMLdrugs, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")


png(filename = paste(figures_folder, '/Fig_07/Panel_E.png', sep = ''), units = "cm", width = 24, height = 24, res = 300)

draw(HMap_BeatAMLdrugs, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")

invisible(dev.off())
```

# 8) Clinical decision tree : Panel F

## - Loading K2 RF clusters

```{r}
RF_cluster_metadata <- read.csv(paste(data_folder, '/RandomForest/Results/Clustering/K2/RF_K2_Clustered_Metadata.csv', 
                                    sep = ''), row.names = 'Patient.ID')
```

## - Data partition

```{r}
Metadata_subset <- RF_cluster_metadata[,c(1:57, 178)]

Metadata_subset$ICC.classification <- factor(as.numeric(factor(Metadata_subset$ICC.classification)))
Metadata_subset$Gender <- factor(Metadata_subset$Gender)
Metadata_subset$Subtype.at.CMML <- factor(Metadata_subset$Subtype.at.CMML)
Metadata_subset$Cutaneous.BPDCN <- factor(Metadata_subset$Cutaneous.BPDCN)
Metadata_subset$RF_Cluster <- factor(Metadata_subset$RF_Cluster)
Metadata_subset$Cytogenetics..at.transformation. <- factor(as.numeric(factor(Metadata_subset$Cytogenetics..at.transformation.)))

set.seed(123)

parts = createDataPartition(Metadata_subset$RF_Cluster, p = .7, list = F)
train_dataset = Metadata_subset[parts, ]
test_dataset = Metadata_subset[-parts, ]
```

## - Feature selection

```{r}
set.seed(123)

boruta_adt <- Boruta(RF_Cluster ~ ., 
                     data = train_dataset, 
                     doTrace = 2, 
                     maxRuns = 3000, 
                     pValue = 0.01, 
                     mcAdj=TRUE)
                     
print(boruta_adt)
boruta_adt <- TentativeRoughFix(boruta_adt)
print(boruta_adt)
plot(boruta_adt, las = 2, cex.axis = 0.7) # cex.axis is used to reduce the font size

confirmed_markers <- names(boruta_adt$finalDecision)[grep(TRUE,boruta_adt$finalDecision == 'Confirmed')]

colMeans(boruta_adt$ImpHistory[,confirmed_markers])

train_dataset_filtered <- train_dataset[,c(confirmed_markers,'RF_Cluster')]
test_dataset_filtered <- test_dataset[,c(confirmed_markers,'RF_Cluster')]

set.seed(123)

# RF modelling

model_filtered = train(RF_Cluster ~ ., data = train_dataset_filtered,
              method = "rf", 
              metric = "ROC",
              trControl = fitControl, 
              tuneLength=10)
model_filtered

# Confirming boruta's features relevance via RF importances

varImp(model_filtered)

set.seed(123)

fitted_filtered <- predict(model_filtered, test_dataset, type = "prob")
fitted_filtered_factor <- factor(ifelse(fitted_filtered$C2 >= 0.60, "C2", "C1"))
```

The RF model is great but is impractical and difficult to implement given that immunophenotype MFIs can be highly variable depending on the antibodies choices and the choice of flow cytometer instrument. We will thus use the percentage of cells positive for each marker as a proxy for the MFI.

## - Loading Immunophenotype as percentage cells positive

```{r}
Immunophenotype_percentage <- read.csv(paste(data_folder, '/Immunophenotype/CHR_BP_CMML_immunophenotype_percentage_positive.csv', 
                                    sep = ''), row.names = 'Patient.ID')

Immunophenotype_percentage <- Immunophenotype_percentage[rownames(RF_cluster_metadata),]
```

## - Decision tree

```{r}
Metadata_subset <- RF_cluster_metadata[,c(1:46,178)]

Metadata_subset$ICC.classification <- factor(as.numeric(factor(Metadata_subset$ICC.classification)))
Metadata_subset$Gender <- factor(Metadata_subset$Gender)
Metadata_subset$Subtype.at.CMML <- factor(Metadata_subset$Subtype.at.CMML)
Metadata_subset$Cutaneous.BPDCN <- factor(Metadata_subset$Cutaneous.BPDCN)
Metadata_subset$RF_Cluster <- factor(Metadata_subset$RF_Cluster)
Metadata_subset$Cytogenetics..at.transformation. <- factor(as.numeric(factor(Metadata_subset$Cytogenetics..at.transformation.)))

Metadata_subset <- cbind(Metadata_subset, round(Immunophenotype_percentage))

set.seed(123)

parts = createDataPartition(Metadata_subset$RF_Cluster, p = .7, list = F)
train_dataset = Metadata_subset[parts, ]
test_dataset = Metadata_subset[-parts, ]

set.seed(123)

train_dataset_filtered <- train_dataset[,c('CD34','CD56','CD117','BM.Flow.percentage','RF_Cluster')]
test_dataset_filtered <- test_dataset[,c('CD34','CD56','CD117','BM.Flow.percentage','RF_Cluster')]

colnames(train_dataset_filtered) <- c('Percent.CD34.pos', 'Percent.CD56.pos', 'Percent.CD117.pos', 'Percent.Blast.BM.Flow','RF_Cluster')
colnames(test_dataset_filtered) <- c('Percent.CD34.pos', 'Percent.CD56.pos', 'Percent.CD117.pos', 'Percent.Blast.BM.Flow','RF_Cluster')

set.seed(123)

treeModel <- rpart::rpart(RF_Cluster ~ ., data = train_dataset_filtered, 
                          control = rpart::rpart.control(minsplit = 3, minbucket = 1, maxdepth = 4,xval = 10,
                                                cp = 0))
rpart::printcp(treeModel)

fit9 <- rpart::prune(treeModel, cp = 0.02)

fitted <- as.data.frame(predict(treeModel, test_dataset_filtered, type = "prob"))
fitted$Identity <- NA

for (patient in rownames(fitted)) {
  if ((fitted[patient,'C1'] > fitted[patient,'C2']) == TRUE) {
  fitted[patient,'Identity'] <- 'C1'
  } else {
  fitted[patient,'Identity'] <- 'C2' 
  }
}

fitted$Identity <- factor(fitted$Identity, levels = c('C1','C2'))

confusionMatrix(data = fitted$Identity, 
                reference = test_dataset$RF_Cluster,
                mode = "everything")

# Plotting decision tree from resulting model

rpart.plot::rpart.plot(treeModel, extra = 2)

png(filename = paste(figures_folder, '/Fig_07/Panel_F.png', sep = ''), units = "cm", width = 18, height = 12, res = 300)

rpart.plot::rpart.plot(treeModel, extra = 2)

invisible(dev.off())
```

## - Evaluating model performance

```{r}
labels <- test_dataset$RF_Cluster

predictions <- fitted$C1
ROCR_pred <- prediction(predictions, labels, label.ordering = c('C2','C1'))
ROCR_pred
ROCR_perf <- performance(ROCR_pred, 'tpr', 'fpr')
plot(ROCR_perf, avg = "threshold", colorize = TRUE)
abline(0, 1, lty = 2)

pROC_pred <- roc(response = labels, predictor = predictions)
plot.roc(pROC_pred, main="pROC")

```

