---
title: "Figure S4"
output: 
  rmdformats::robobook:
    lightbox: true
    fig_caption: true
---

<style type="text/css">
  body{
  font-size: 10pt;
}
</style>

# 1) Setup

## - Defining work directory

In this section, we define the working directory for the R Markdown document.

```{r setup, include=TRUE}
# Get the path of the current script
# Then get the parent directory of the parent directory of the parent directory
local_wd_folder <- dirname(dirname(dirname(rstudioapi::getSourceEditorContext()$path)))

# Set the root directory for knitr to the local working directory
knitr::opts_knit$set(root.dir = local_wd_folder)
```

## - Defining input data and output directories

Here, we define the directories for input data and output files.

```{r}
# Get the directory of the current script
script_folder <- dirname(rstudioapi::getSourceEditorContext()$path)

# Define the data folder and output folder
data_folder <- './Data'
#results_folder <- './Results'
figures_folder <- './Figures'
```

## - Setting seed

Setting a seed ensures that any random processes are reproducible.

```{r}
# Set a seed for reproducibility
set.seed(123)
```

## - Packages installation (optional)

In this section, we ensure that all necessary packages are installed.

```{r}
# Ensure BiocManager is available for installation of Bioconductor packages
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
BiocManager::install(version = "3.18")

# Define a list of required packages used in this script
packages_required <- c("escape")

# Identify any required packages that are not installed
packages_uninstalled <- packages_required[!(packages_required %in% installed.packages()[,"Package"])]

# Install any uninstalled packages
if(length(packages_uninstalled)) BiocManager::install(packages_uninstalled)
```

## - Loading packages

Here, we load the necessary packages for our analysis.

```{r message=FALSE, warning=FALSE}
# Load stringr for string manipulation
library(escape, quietly = TRUE)
```

## - Loading palettes

In this section, we load additional color palettes and define some custom ones.

```{r message=FALSE, warning=FALSE}
# Load additional colour palette packages
library(unikn, quietly = TRUE)
library(RColorBrewer, quietly = TRUE)
library(yarrr, quietly = TRUE)
library(scales, quietly = TRUE)
library(ggsci, quietly = TRUE)

# Define a set of custom color palettes from the unikn package
mix_1 <- usecol(pal = c(Karpfenblau, "white", Peach), n = 15)
mix_2 <- usecol(pal = c(rev(pal_seeblau), "white", pal_pinky))
mix_3 <- usecol(pal = c(rev(pal_bordeaux), "white", pal_petrol), n = 15)

# Display the custom palettes
seecol(list(mix_1, mix_2, mix_3), col_brd = "white", lwd_brd = 4, title = "Comparing palettes mixed from unikn colors", pal_names = c("mix_1", "mix_2", "mix_3"))

# Define a second set of custom palettes from the RColorBrewer and yarrr packages
brew_mix <- usecol(c(rev(brewer.pal(n = 4, name = "Reds")), "white", brewer.pal(n = 4, name = "Blues")), n = 13)
brew_ext <- usecol(brewer.pal(n = 11, name = "Spectral"), n = 12)
yarrr_mix <- usecol(c(piratepal("nemo"), piratepal("bugs")))
yarrr_mod <- usecol(c(piratepal("ipod")), n = 9)

# Display the second set of custom palettes
seecol(pal = list(brew_mix, brew_ext, yarrr_mix, yarrr_mod), col_brd = "white", lwd_brd = 2, title = "Using usecol() and seecol() to mix and modify palettes", pal_names = c("brew_mix", "brew_ext", "yarrr_mix", "yarrr_mod"))

# Define additional custom palettes from the scales package
natjournals_palette <- pal_npg("nrc")(10)
```

## - Log Session Info

Finally, we log the session information for reproducibility.

```{r message=TRUE, warning=TRUE}
# Write the session information to a text file
writeLines(capture.output(sessionInfo()), file.path(script_folder, 'Fig_S4_SessionInfo.txt'))

# Print the session information
sessionInfo()
```

# 2) Loading input files

```{r}
Expression_matrix <- read.csv(paste(data_folder, '/RNASeq/InHouse/Counts/Processed_counts/Grouped/Tximport_TPM_mat.csv', 
                                    sep = ''), 
                              row.names = 'X')

colnames(Expression_matrix) <- str_replace_all(colnames(Expression_matrix), pattern = 'X', replacement = '')

Genes_annotation_metadata <- read.csv(paste(data_folder, "/Resources/RNASeq/Annotation_files/GENCODE_v42.gene_metadata.csv", sep = ''))
rownames(Genes_annotation_metadata) <- Genes_annotation_metadata$gene_id

Samples_metadata <- read.csv(paste(data_folder, '/RNASeq/InHouse/Metadata/Samples_Metadata.csv', sep = ''), row.names = 'Patient.ID')
```

# 3) Ordering expression matrix (TPM)

```{r}
Unique_genes <- read.table(file = paste(data_folder, '/RNASeq/InHouse/Counts/Processed_counts/List_unique_genes.csv', 
                                    sep = ''), header = T, sep = ',')$unique_genes

Expression_matrix <- Expression_matrix[Unique_genes, ]

rownames(Expression_matrix) <- Genes_annotation_metadata[rownames(Expression_matrix),'gene_name']

Expression_matrix <- as.matrix(Expression_matrix)

Expression_matrix <- Expression_matrix[,rownames(Samples_metadata)]
```

# 4) Creating folders for Pathway Analyses

```{r}
dir.create(file.path(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis', 
                                    sep = '')), showWarnings = FALSE)

dir.create(file.path(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments', 
                                    sep = '')), showWarnings = FALSE)

dir.create(file.path(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/CellType', 
                                    sep = '')), showWarnings = FALSE)

dir.create(file.path(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Resistance', 
                                    sep = '')), showWarnings = FALSE)

dir.create(file.path(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Response', 
                                    sep = '')), showWarnings = FALSE)

dir.create(file.path(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/WGCNA', 
                                    sep = '')), showWarnings = FALSE)

dir.create(file.path(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/Control_vs_BP-CMML', 
                                    sep = '')), showWarnings = FALSE)
```

# 5) Gene signature

## - CellTypes

### - 2019 VanGalen P. V. et al AML signatures

```{r}
VanGalen_CellType_Modules  <- read.csv(paste(data_folder, '/Resources/RNASeq/Gene_Pathways/2019__VanGalen_et_al__AML_signatures.csv', 
                                    sep = ''))

HSC.like <- GSEABase::GeneSet(VanGalen_CellType_Modules$HSC.like)
HSC.like <- GSEABase::`setName<-`(object = HSC.like, value='HSC.like')

Progenitor.like <- GSEABase::GeneSet(VanGalen_CellType_Modules$Progenitor.like)
Progenitor.like <- GSEABase::`setName<-`(object = Progenitor.like, value='Progenitor.like')

HSC.Prog.like <- GSEABase::GeneSet(VanGalen_CellType_Modules$HSC.Prog.like)
HSC.Prog.like <- GSEABase::`setName<-`(object = HSC.Prog.like, value='HSC.Prog.like')

GMP.like <- GSEABase::GeneSet(VanGalen_CellType_Modules$GMP.like)
GMP.like <- GSEABase::`setName<-`(object = GMP.like, value='GMP.like')

Promono.like <- GSEABase::GeneSet(VanGalen_CellType_Modules$Promono.like)
Promono.like <- GSEABase::`setName<-`(object = Promono.like, value='Promono.like')

Monocyte.like <- GSEABase::GeneSet(VanGalen_CellType_Modules$Monocyte.like)
Monocyte.like <- GSEABase::`setName<-`(object = Monocyte.like, value='Monocyte.like')

cDC.like <- GSEABase::GeneSet(VanGalen_CellType_Modules$cDC.like)
cDC.like <- GSEABase::`setName<-`(object = cDC.like, value='cDC.like')

Myeloid.like <- GSEABase::GeneSet(VanGalen_CellType_Modules$Myeloid.like)
Myeloid.like <- GSEABase::`setName<-`(object = Myeloid.like, value='Myeloid.like')

VanGalen_CellType_Modules_collection <- GSEABase::GeneSetCollection(HSC.like, HSC.Prog.like, Progenitor.like, 
                                                            GMP.like, Promono.like, Monocyte.like, 
                                                            cDC.like, Myeloid.like)

# - - - - Scoring samples excluding healthies

VanGalen_CellType_Modules_scaled_results_no_healthies <- enrichIt(obj = as.matrix(Expression_matrix[,rownames(Samples_metadata[Samples_metadata$Condition == 'BP-CMML',])]), 
               gene.sets = VanGalen_CellType_Modules_collection, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)

write.csv(VanGalen_CellType_Modules_scaled_results_no_healthies, paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/CellType/2019__VanGalen_et_al__AML_signatures__Enrichments.csv', 
                                    sep = ''),
          quote = F, row.names = T)
```

### - 2017 Velten L. et al Healthy signatures

```{r}
Velten_CellType_Modules <- read.csv(paste(data_folder, '/Resources/RNASeq/Gene_Pathways/2017__Velten_et_al__Healthy_signatures.csv', 
                                    sep = ''))

library(stringi)

Hematopoietic_module_genes <- list()

for (Hematopoietic_module in colnames(Velten_CellType_Modules)) {

Hematopoietic_module_genes[[Hematopoietic_module]] <-  Velten_CellType_Modules[[Hematopoietic_module]][Velten_CellType_Modules[[Hematopoietic_module]] != ""]
  
Hematopoietic_module_genes[[Hematopoietic_module]] <- str_split_fixed(Hematopoietic_module_genes[[Hematopoietic_module]], pattern = ' ', n = 2)[,1]

Hematopoietic_module_genes[[Hematopoietic_module]] <- as.vector(na.omit(Hematopoietic_module_genes[[Hematopoietic_module]]))

Hematopoietic_module_genes[[Hematopoietic_module]] <- GSEABase::GeneSet(Hematopoietic_module_genes[[Hematopoietic_module]])

Hematopoietic_module_genes[[Hematopoietic_module]] <- GSEABase::`setName<-`(object = Hematopoietic_module_genes[[Hematopoietic_module]], value = sprintf('%s',Hematopoietic_module))

}

Velten_Celltype_Modules <- GSEABase::GeneSetCollection(Hematopoietic_module_genes)

# - - - - Scoring samples excluding healthies

Velten_Celltype_Modules_scaled_results_no_healthies <- enrichIt(obj = as.matrix(Expression_matrix[,rownames(Samples_metadata[Samples_metadata$Condition == 'BP-CMML',])]), 
               gene.sets = Velten_Celltype_Modules, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)

write.csv(Velten_Celltype_Modules_scaled_results_no_healthies, paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/CellType/2017__Velten_et_al__Healthy_signatures__Enrichments.csv', 
                                    sep = ''),
          quote = F, row.names = T)
```

## - Drug Resistance

### - 2017 Unnikrishnan A. et al Azacytidine resistance/response signatures

```{r}
AZA_modules <- read.csv(paste(data_folder, '/Resources/RNASeq/Gene_Pathways/2017__Unnikrishnan_et_al_MDS_expression_preAZA_treatment.csv', 
                                    sep = ''))

AZA_modules_UP <- AZA_modules %>% dplyr::filter(Comment == 'UP in AZA responders')
AZA_modules_UP <- GSEABase::GeneSet(AZA_modules_UP$Gene.Symbol..)
AZA_modules_UP <- GSEABase::`setName<-`(object = AZA_modules_UP, value='AZA response')

AZA_modules_DOWN <- AZA_modules %>% dplyr::filter(Comment == 'UP in AZA non-responders')
AZA_modules_DOWN <- GSEABase::GeneSet(AZA_modules_DOWN$Gene.Symbol..)
AZA_modules_DOWN <- GSEABase::`setName<-`(object = AZA_modules_DOWN, value='AZA resistance')

AZA_modules_collection <- GSEABase::GeneSetCollection(AZA_modules_UP, AZA_modules_DOWN)

# - - - - Scoring samples excluding healthies

AZA_modules_scaled_result_no_healthies <- enrichIt(obj = as.matrix(Expression_matrix[,rownames(Samples_metadata[Samples_metadata$Condition == 'BP-CMML',])]), 
               gene.sets = AZA_modules_collection, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)

write.csv(AZA_modules_scaled_result_no_healthies %>% select(AZA.resistance), 
          paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Resistance/2017__Unnikrishnan_et_al__AZA_resistance_signature__Enrichments.csv', 
                                    sep = ''),
          quote = F, row.names = T)

write.csv(AZA_modules_scaled_result_no_healthies %>% select(AZA.response), 
          paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Response/2017__Unnikrishnan_et_al__AZA_response_signature__Enrichments.csv', 
                                    sep = ''),
          quote = F, row.names = T)

# - - - - Scoring samples including healthies

AZA_modules_scaled_result_with_healthies <- enrichIt(obj = as.matrix(Expression_matrix), 
               gene.sets = AZA_modules_collection, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)

write.csv(AZA_modules_scaled_result_with_healthies %>% select(AZA.resistance), 
          paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Resistance/2017__Unnikrishnan_et_al__AZA_resistance_signature__Enrichments_w_healthies.csv', 
                                    sep = ''),
          quote = F, row.names = T)

write.csv(AZA_modules_scaled_result_with_healthies %>% select(AZA.response), 
          paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Response/2017__Unnikrishnan_et_al__AZA_response_signature__Enrichments_w_healthies.csv', 
                                    sep = ''),
          quote = F, row.names = T)
```

### - 2020 Zhang H et al. Venetoclax resistance signature

```{r}
Veneto_resistance_module <- read.csv(paste(data_folder, '/Resources/RNASeq/Gene_Pathways/2020__Zhang_et_al__venetoclax_resistance_signature.csv', 
                                    sep = ''))

Veneto_resistance_module <- Veneto_resistance_module %>% filter(fold_change > 0)

Veneto_resistance_module$Rank <- 1-percent_rank(Veneto_resistance_module$fdr)

Veneto_resistance_module <- Veneto_resistance_module %>% arrange(fdr)

Veneto_resistance_module_top_10percent_genes <- Veneto_resistance_module[grep(TRUE,percent_rank(Veneto_resistance_module$Rank) >= 0.90),'symbol']
Veneto_resistance_module_genes <- GSEABase::GeneSet(Veneto_resistance_module_top_10percent_genes)
Veneto_resistance_module_genes <- GSEABase::`setName<-`(object = Veneto_resistance_module_genes, value='Venetoclax resistance')
Veneto_resistance_module_genes <- GSEABase::GeneSetCollection(Veneto_resistance_module_genes)

# - - - - Scoring samples excluding healthies

Veneto_resistance_module_scaled_result_no_healthies <- enrichIt(obj = as.matrix(Expression_matrix[,rownames(Samples_metadata[Samples_metadata$Condition == 'BP-CMML',])]), 
               gene.sets = Veneto_resistance_module_genes, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)

write.csv(Veneto_resistance_module_scaled_result_no_healthies, 
          paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Resistance/2020__Zhang_et_al__Ven_resistance_signature__Enrichments.csv', 
                                    sep = ''),
          quote = F, row.names = T)

# - - - - Scoring samples including healthies

Veneto_resistance_module_scaled_result_with_healthies <- enrichIt(obj = as.matrix(Expression_matrix), 
               gene.sets = Veneto_resistance_module_genes, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)

write.csv(Veneto_resistance_module_scaled_result_with_healthies, 
          paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Resistance/2020__Zhang_et_al__Ven_resistance_signature__Enrichments_w_healthies.csv', 
                                    sep = ''),
          quote = F, row.names = T)
```

### - 2020 Williams M. et al. Daunorubicin resistance signature

```{r}
Daunorubicin_resistance_module <- read.csv(paste(data_folder, '/Resources/RNASeq/Gene_Pathways/2020__Williams_et_al__daunorubicin_resistance_signature.csv', 
                                    sep = ''))

Daunorubicin_resistance_module$Log2FC <- log2(Daunorubicin_resistance_module$FC)

Daunorubicin_resistance_module <- Daunorubicin_resistance_module %>% filter(Log2FC > 0)

Daunorubicin_resistance_module_genes <- GSEABase::GeneSet(Daunorubicin_resistance_module$symbol[which((Daunorubicin_resistance_module$symbol == '') == F)])

Daunorubicin_resistance_module_genes <- GSEABase::`setName<-`(object = Daunorubicin_resistance_module_genes, value='Daunorubicin resistance')

Daunorubicin_resistance_module_genes <- GSEABase::GeneSetCollection(Daunorubicin_resistance_module_genes)

# - - - - Scoring samples excluding healthies

Daunorubicin_resistance_module_scaled_result_no_healthies <- enrichIt(obj = as.matrix(Expression_matrix[,rownames(Samples_metadata[Samples_metadata$Condition == 'BP-CMML',])]), 
               gene.sets = Daunorubicin_resistance_module_genes, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)

write.csv(Daunorubicin_resistance_module_scaled_result_no_healthies, 
          paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Resistance/2020__Williams_et_al__Daunorubicin_resistance_signature__Enrichments.csv', 
                                    sep = ''),
          quote = F, row.names = T)

# - - - - Scoring samples including healthies

Daunorubicin_resistance_module_scaled_result_with_healthies <- enrichIt(obj = as.matrix(Expression_matrix), 
               gene.sets = Daunorubicin_resistance_module_genes, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)

write.csv(Daunorubicin_resistance_module_scaled_result_with_healthies, 
          paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Resistance/2020__Williams_et_al__Daunorubicin_resistance_signature__Enrichments_w_healthies.csv', 
                                    sep = ''),
          quote = F, row.names = T)
```

## - Drug Response

### - 2019 Xu H. et al. Cytarabine response signature

```{r}
cytarabine_response_module <- read.csv(paste(data_folder, '/Resources/RNASeq/Gene_Pathways/2019__Xu_et_al_cytarabine_response_signature.csv', 
                                    sep = ''))

cytarabine_response_module <- cytarabine_response_module %>% filter(Log2_FoldChange > 0)

cytarabine_response_module_genes <- GSEABase::GeneSet(cytarabine_response_module$GeneName)
cytarabine_response_module_genes <- GSEABase::`setName<-`(object = cytarabine_response_module_genes, value='Cytarabine response')
cytarabine_response_module_genes <- GSEABase::GeneSetCollection(cytarabine_response_module_genes)

# - - - - Scoring samples excluding healthies

cytarabine_response_module_scaled_result_no_healthies <- enrichIt(obj = as.matrix(Expression_matrix[,rownames(Samples_metadata[Samples_metadata$Condition == 'BP-CMML',])]), 
               gene.sets = cytarabine_response_module_genes, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)

write.csv(cytarabine_response_module_scaled_result_no_healthies, 
          paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Response/2019__Xu_et_al_cytarabine_response_signature__Enrichments.csv', 
                                    sep = ''),
          quote = F, row.names = T)

# - - - - Scoring samples including healthies

cytarabine_response_module_scaled_result_with_healthies <- enrichIt(obj = as.matrix(Expression_matrix), 
               gene.sets = cytarabine_response_module_genes, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)

write.csv(cytarabine_response_module_scaled_result_with_healthies, 
          paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Response/2019__Xu_et_al_cytarabine_response_signature__Enrichments_w_healthies.csv', 
                                    sep = ''),
          quote = F, row.names = T)
```

### - 2022 Vizome BEAT AML mined drug response signatures

```{r}
Finaldrugs_selection <- read.csv(paste(data_folder, '/RNASeq/PubliclyAvailable/2022__Vizome__BeatAML/Sample_Selection/Drugs_passing_threshold.csv', 
                                    sep = ''))$Drug

DrugResponses <- list()
DrugResponses_genes <- list()

for (drug in Finaldrugs_selection) {
  
  DrugResponses[[drug]] <- read.csv(paste(data_folder, '/RNASeq/PubliclyAvailable/2022__Vizome__BeatAML/Mined_Signatures/', 
                                          drug, '/',
                                          sprintf("%s_Responsive_vs_Resistant_downregulated_padj_significant_DEGs_LFC_Ashr.csv", drug),
                                    sep = ''))
  
  DrugResponses[[drug]]$rank <- NULL
  
  if ((length(DrugResponses[[drug]]$symbol) < 10) == TRUE) {
  DrugResponses[[drug]] <- NULL 
  } else {
    if ((length(DrugResponses[[drug]]$symbol) > 500) == TRUE) {
        DrugResponses[[drug]]$Rank <- 1-percent_rank(DrugResponses[[drug]]$padj)
        DrugResponses_genes[[drug]] <- DrugResponses[[drug]][grep(TRUE,percent_rank(DrugResponses[[drug]]$Rank) >= 0.70),'symbol']
          if ((length(DrugResponses_genes[[drug]]) > 500) == TRUE) {
            DrugResponses_genes[[drug]] <- GSEABase::GeneSet(na.omit(unique(DrugResponses[[drug]]$symbol[1:500])))
            DrugResponses_genes[[drug]] <- GSEABase::`setName<-`(object = DrugResponses_genes[[drug]], value=sprintf('%s',drug)) 
          }
          else {
            DrugResponses_genes[[drug]] <- GSEABase::GeneSet(na.omit(unique(DrugResponses_genes[[drug]])))
            DrugResponses_genes[[drug]] <- GSEABase::`setName<-`(object = DrugResponses_genes[[drug]], value=sprintf('%s',drug)) 
          }
        }
    else {
        DrugResponses_genes[[drug]] <- GSEABase::GeneSet(na.omit(unique(DrugResponses[[drug]]$symbol[1:500])))
        DrugResponses_genes[[drug]] <- GSEABase::`setName<-`(object = DrugResponses_genes[[drug]], value=sprintf('%s',drug)) 
      }
    }
  }

  df <- data.frame(row.names = names(DrugResponses_genes))
  for (drug in names(DrugResponses_genes)) {
  df[drug,'length'] <- length(GSEABase::geneIds(DrugResponses_genes[[drug]]))
  }
  
  DrugResponseSignatures <- GSEABase::GeneSetCollection(DrugResponses_genes)

# - - - - Scoring samples excluding healthies
  
  DrugResponseSignatures_scaled_results_no_healthies <- enrichIt(obj = as.matrix(Expression_matrix[,rownames(Samples_metadata[Samples_metadata$Condition == 'BP-CMML',])]), 
               gene.sets = DrugResponseSignatures, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)

  colnames(DrugResponseSignatures_scaled_results_no_healthies) <- gsub("\\.{2}", " ", colnames(DrugResponseSignatures_scaled_results_no_healthies))
  colnames(DrugResponseSignatures_scaled_results_no_healthies) <- gsub("\\.*$","", colnames(DrugResponseSignatures_scaled_results_no_healthies))
  DrugResponseSignatures_scaled_results_no_healthies <- DrugResponseSignatures_scaled_results_no_healthies[,sort(colnames(DrugResponseSignatures_scaled_results_no_healthies))]
  
  write.csv(DrugResponseSignatures_scaled_results_no_healthies, 
            paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Response/2022__BeatAMLDrugResponseSignatures__Enrichments.csv', 
                                      sep = ''),
            quote = F, row.names = T)
```

## - WGCNA

```{r}
WGCNA_top_genes <- read.csv(paste(data_folder, '/RNASeq/InHouse/WGCNA/WGCNA_Modules_Top10_percent_genes.csv', 
                                    sep = ''))

colnames(WGCNA_top_genes) <- c('Module', 'Gene')

Top_ME1 <- WGCNA_top_genes %>% dplyr::filter(Module == 'ME1')
Top_ME1 <- GSEABase::GeneSet(Top_ME1$Gene)
Top_ME1 <- GSEABase::`setName<-`(object = Top_ME1, value='ME1')

Top_ME2 <- WGCNA_top_genes %>% dplyr::filter(Module == 'ME2')
Top_ME2 <- GSEABase::GeneSet(Top_ME2$Gene)
Top_ME2<- GSEABase::`setName<-`(object = Top_ME2, value='ME2')

Top_ME3 <- WGCNA_top_genes %>% dplyr::filter(Module == 'ME3')
Top_ME3 <- GSEABase::GeneSet(Top_ME3$Gene)
Top_ME3 <- GSEABase::`setName<-`(object = Top_ME3, value='ME3')

Top_ME4 <- WGCNA_top_genes %>% dplyr::filter(Module == 'ME4')
Top_ME4 <- GSEABase::GeneSet(Top_ME4$Gene)
Top_ME4 <- GSEABase::`setName<-`(object = Top_ME4, value='ME4')

Top_ME5 <- WGCNA_top_genes %>% dplyr::filter(Module == 'ME5')
Top_ME5 <- GSEABase::GeneSet(Top_ME5$Gene)
Top_ME5 <- GSEABase::`setName<-`(object = Top_ME5, value='ME5')

Top_ME6 <- WGCNA_top_genes %>% dplyr::filter(Module == 'ME6')
Top_ME6 <- GSEABase::GeneSet(Top_ME6$Gene)
Top_ME6 <- GSEABase::`setName<-`(object = Top_ME6, value='ME6')

WGCNA_modules_top_genes <- GSEABase::GeneSetCollection(Top_ME1, 
                                                       Top_ME2,
                                                       Top_ME3,
                                                       Top_ME4,
                                                       Top_ME5,
                                                       Top_ME6)

# - - - - Scoring samples excluding healthies
  
  WGCNA_scaled_score_no_healthies <- enrichIt(obj = as.matrix(Expression_matrix[,rownames(Samples_metadata[Samples_metadata$Condition == 'BP-CMML',])]), 
               gene.sets = WGCNA_modules_top_genes, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)
  
  write.csv(WGCNA_scaled_score_no_healthies, 
            paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/WGCNA/WGCNA_modules__Enrichments.csv', 
                                      sep = ''),
            quote = F, row.names = T)
  
# - - - - Scoring samples including healthies
  
  WGCNA_scaled_score_with_healthies <- enrichIt(obj = as.matrix(Expression_matrix), 
               gene.sets = WGCNA_modules_top_genes, 
               cores = 12, 
               ssGSEA.norm = T,
               min.size = 5)
  
  write.csv(WGCNA_scaled_score_with_healthies, 
            paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/WGCNA/WGCNA_modules__Enrichments_w_healthies.csv', 
                                      sep = ''),
            quote = F, row.names = T)
```

# 6) Plotting

## - 2019 VanGalen P. V. et al. Enrichment Scores HeatMap

```{r}
VanGalen_celltype_modules_scores <- read.csv(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/CellType/2019__VanGalen_et_al__AML_signatures__Enrichments.csv', 
                                    sep = ''), row.names = 'X')

colnames(VanGalen_celltype_modules_scores) <- c('HSC-like',
                                     'HSC.Prog-like',
                                     'Progenitor-like',
                                     'GMP-like',
                                     'Promono-like',
                                     'Monocyte-like',
                                     'cDC-like',
                                     'Myeloid-like')

VanGalen_celltype_modules_scores <- t(VanGalen_celltype_modules_scores)

Identity <- data.frame(row.names = rownames(VanGalen_celltype_modules_scores))

breaks_val <- seq(from = 0, to = 1, length.out = 14)
col_heatmap_guide <- colorRamp2(breaks = breaks_val,colors = rev(c("#67000D",brew_mix)))

HMap_vangalen <- Heatmap(VanGalen_celltype_modules_scores, 
        name = 'ssGSEA Score',
        col = col_heatmap_guide, rect_gp = gpar(col = "black", lwd = 0.05),
        column_title_gp = gpar(fontsize=11, fontface='bold'),
        row_names_gp = gpar(fontsize=10), cluster_rows = F,
        cluster_column_slices = F,
        column_names_gp = gpar(fontsize=10), show_heatmap_legend = F,
        heatmap_legend_param = list(legend_direction = "horizontal",
                                    legend_height = unit(5, "cm")),
        row_names_side = 'right')

draw(HMap_vangalen, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")

# Open a new PNG device for plotting
png(filename = paste(figures_folder, '/Supplemental/Fig_S4-1.png', sep = ''), units = 'cm', width = 18.62, height = 7, res = 300)

draw(HMap_vangalen, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")

invisible(dev.off())
```

## - 2017 Velten L. et al Enrichment Scores HeatMap

```{r}
Velten_celltype_modules_scores <- read.csv(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/CellType/2017__Velten_et_al__Healthy_signatures__Enrichments.csv', 
                                    sep = ''), row.names = 'X')

colnames(Velten_celltype_modules_scores) <- c('HOXA3.HOXB6',
                               'HLF.ZFP36L2',
                               'SPINK2.SELL',
                               'FLT3.SATB1',
                               'IRF1.CASP1',
                               'EBF1.ID3',
                               'EAF2.KLF4',
                               'IRF7.IRF8',
                               'SPI1.GFI1',
                               'CEBPA.CEBPD',
                               'HDC.LMO4',
                               'GATA2.NFE2',
                               'TAL1.HSF1',
                               'GP1BB.PBX1',
                               'GATA1.KLF1')

Velten_celltype_modules_scores <- t(Velten_celltype_modules_scores)

Module_Identity <- data.frame(row.names = rownames(Velten_celltype_modules_scores))

Module_Identity[c('HOXA3.HOXB6',
                  'HLF.ZFP36L2',
                  'SPINK2.SELL'),'Celltype'] <- 'HSC/MPP'
Module_Identity[c('FLT3.SATB1'),'Celltype'] <- 'LMPP'
Module_Identity[c('IRF1.CASP1',
                  'EBF1.ID3'),'Celltype'] <- 'CLP'
Module_Identity[c('EAF2.KLF4',
                  'IRF7.IRF8'),'Celltype'] <- 'MDP'
Module_Identity[c('SPI1.GFI1',
                  'CEBPA.CEBPD'),'Celltype'] <- 'NeP'
Module_Identity[c('HDC.LMO4'),'Celltype'] <- 'EoBaMaP'
Module_Identity[c('GATA2.NFE2',
                  'TAL1.HSF1'),'Celltype'] <- 'MEP'
Module_Identity[c('GP1BB.PBX1'),'Celltype'] <- 'MkP'
Module_Identity[c('GATA1.KLF1'),'Celltype'] <- 'EP'

Module_Identity$Celltype <- factor(Module_Identity$Celltype, 
                                   levels = c('HSC/MPP',
                                              'LMPP',
                                              'CLP',
                                              'MDP',
                                              'NeP',
                                              'EoBaMaP',
                                              'MEP',
                                              'MkP',
                                              'EP'))

Celltype <- data.frame(row.names = c('HSC/MPP',
                                              'LMPP',
                                              'CLP',
                                              'MDP',
                                              'NeP',
                                              'EoBaMaP',
                                              'MEP',
                                              'MkP',
                                              'EP'))


Celltype$Colour <- c('#000000',
                         '#000000',
                         '#000000',
                         '#000000',
                         '#000000',
                         '#000000',
                         '#000000',
                         '#000000',
                         '#000000')

breaks_val <- seq(from = 0, to = 1, length.out = 14)
col_heatmap_guide <- colorRamp2(breaks = breaks_val,colors = rev(c("#67000D",brew_mix)))

HMap <- Heatmap(Velten_celltype_modules_scores, 
        name = 'ssGSEA Score',
        col = col_heatmap_guide, rect_gp = gpar(col = "black", lwd = 0.05),
        column_title_gp = gpar(fontsize=11, fontface='bold'),
        row_split = Module_Identity$Celltype, cluster_rows = F,
        row_gap = unit(0.05, "cm"), 
        cluster_column_slices = F,
        row_names_gp = gpar(fontsize=9), row_title_rot = 0,
        row_title_gp = gpar(fontsize=9, fontface='bold', col=Celltype$Colour),
        column_names_gp = gpar(fontsize=9), show_heatmap_legend = F,
        heatmap_legend_param = list(legend_direction = "horizontal",
                                    legend_height = unit(5, "cm")),
        row_names_side = 'right')

draw(HMap, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")

# Open a new PNG device for plotting
png(filename = paste(figures_folder, '/Supplemental/Fig_S4-2.png', sep = ''), units = 'cm', width = 19.2, height = 11, res = 300)

draw(HMap, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")

invisible(dev.off())
```

## - 2022 Vizome BeatAML drug response scores Heatmap

```{r}
BeatAML_drug_response_modules_scores <- read.csv(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Response/2022__BeatAMLDrugResponseSignatures__Enrichments.csv', 
                                    sep = ''), row.names = 'X')

# We excluded common drugs for which we have curated signatures

BeatAML_drug_response_modules_scores <- BeatAML_drug_response_modules_scores[,setdiff(colnames(BeatAML_drug_response_modules_scores),
                                                                                      c('Azacytidine', 'Venetoclax', 'Cytarabine'))]

colnames(BeatAML_drug_response_modules_scores) <- c(
                                "A-674563",
                                "ABT-737",
                                "Afatinib (BIBW-2992)",
                                "AT7519",
                                "Axitinib (AG-013736)",
                                "AZD1152-HQPA (AZD2811)",
                                "AZD1480",
                                "Bay 11-7085",
                                "BEZ235",
                                "Birinapant",
                                "Bortezomib (Velcade)",
                                "Bosutinib (SKI-606)",
                                "Cabozantinib",
                                "CI-1040 (PD184352)",
                                "Crenolanib",
                                "Crizotinib (PF-2341066)",
                                "CYT387",
                                "Dasatinib",
                                "Doramapimod (BIRB 796)",
                                "Dovitinib (CHIR-258)",
                                "Elesclomol",
                                "Entospletinib (GS-9973)",
                                "Entrectinib",
                                "Flavopiridol",
                                "Foretinib (XL880)",
                                "GDC-0941",
                                "Gilteritinib",
                                "GSK-1838705A",
                                "GW-2580",
                                "H-89",
                                "Ibrutinib (PCI-32765)",
                                "Idelalisib",
                                "Indisulam",
                                "INK-128",
                                "JAK Inhibitor I",
                                "JNJ-28312141",
                                "JNJ-38877605",
                                "JQ1",
                                "KI20227",
                                "KU-55933",
                                "KW-2449",
                                "Lenalidomide",
                                "Lenvatinib",
                                "Linifanib (ABT-869)",
                                "Masitinib (AB-1010)",
                                "MGCD-265",
                                "Midostaurin",
                                "MK-2206",
                                "MLN120B",
                                "Neratinib (HKI-272)",
                                "NF-kB Activation Inhibitor",
                                "Nutlin 3a",
                                "NVP-ADW742",
                                "NVP-TAE684",
                                "OTX-015",
                                "Panobinostat",
                                "PD173955",
                                "Pelitinib (EKB-569)",
                                "Perhexiline maleate",
                                "PH-797804",
                                "PHA-665752",
                                "PI-103",
                                "Ponatinib (AP24534)",
                                "PP242",
                                "PRT062607",
                                "Quizartinib (AC220)",
                                "RAF265 (CHIR-265)",
                                "Ralimetinib (LY2228820)",
                                "Ranolazine",
                                "Rapamycin",
                                "Regorafenib (BAY 73-4506)",
                                "S31-201",
                                "Selinexor",
                                "Selumetinib (AZD6244)",
                                "SNS-032 (BMS-387032)",
                                "Sorafenib",
                                "Sunitinib",
                                "Tivozanib (AV-951)",
                                "Tofacitinib (CP-690550)",
                                "Tozasertib (VX-680)",
                                "Trametinib (GSK1120212)",
                                "Vandetanib (ZD6474)",
                                "Vargetef",
                                "Vatalanib (PTK787)",
                                "Volasertib (BI-6727)",
                                "VX-745",
                                "Tanespimycin (17-AAG)")

breaks_val <- seq(from = 0, to = 1, length.out = 14)
col_heatmap_guide <- colorRamp2(breaks = breaks_val,colors = rev(c("#67000D",brew_mix)))

HMap_BeatAMLdrugs <- Heatmap(BeatAML_drug_response_modules_scores, 
        name = 'ssGSEA Score',
        col = col_heatmap_guide, 
        rect_gp = gpar(col = "black", lwd = 0.05),
        cluster_row_slices = F,
        column_title_gp = gpar(fontsize=14, fontface='bold'),
        row_names_gp = gpar(fontsize=8), 
        cluster_rows = T,
        column_names_gp = gpar(fontsize=8), 
        show_heatmap_legend = T,
        heatmap_legend_param = list(legend_direction = "horizontal",
                                    legend_width = unit(2, "cm")),
        row_names_side = 'right')

draw(HMap_BeatAMLdrugs, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")

# Open a new PNG device for plotting
png(filename = paste(figures_folder, '/Supplemental/Fig_S4-3.png', sep = ''), units = 'cm', width = 30, height = 20, res = 300)

draw(HMap_BeatAMLdrugs, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")

dev.off()
```

