---
title: "Figure 06 - Panels DEF"
output: 
  rmdformats::robobook:
    lightbox: true
    fig_caption: true
---

```{=html}
<style type="text/css">
  body{
  font-size: 10pt;
}
</style>
```

# 1) Setup

## - Defining work directory

In this section, we define the working directory for the R Markdown document.

```{r setup, include=TRUE}
# Get the path of the current script
# Then get the parent directory of the parent directory of the parent directory
local_wd_folder <- dirname(dirname(dirname(rstudioapi::getSourceEditorContext()$path)))

# Set the root directory for knitr to the local working directory
knitr::opts_knit$set(root.dir = local_wd_folder)
```

## - Defining input data and output directories

Here, we define the directories for input data and output files.

```{r}
# Get the directory of the current script
script_folder <- dirname(rstudioapi::getSourceEditorContext()$path)

# Define the data folder and output folder
data_folder <- './Data'
#results_folder <- './Results'
figures_folder <- './Figures'

dir.create(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/Pathway_Correlations', sep = ''), 
           showWarnings = F)
```

## - Setting seed

Setting a seed ensures that any random processes are reproducible.

```{r}
# Set a seed for reproducibility
set.seed(123)
```

## - Packages installation (optional)

In this section, we ensure that all necessary packages are installed.

```{r}
# Ensure BiocManager is available for installation of Bioconductor packages
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
BiocManager::install(version = "3.18", ask = FALSE)

# Define a list of required packages used in this script
packages_required <- c("ggplot2", "dplyr", "smplot2", "WGCNA", "tidyr",
                       "ggrepel", "ComplexHeatmap",
                       "unikn", "RColorBrewer", "yarrr", "scales", "ggsci")

# Identify any required packages that are not installed
packages_uninstalled <- packages_required[!(packages_required %in% installed.packages()[,"Package"])]

# Install any uninstalled packages
if(length(packages_uninstalled)) BiocManager::install(packages_uninstalled, ask = FALSE)
```

## - Loading packages

Here, we load the necessary packages for our analysis.

```{r message=FALSE, warning=FALSE}
library(ggplot2, quietly = TRUE)
library(ggrepel, quietly = TRUE)
library(smplot2, quietly = TRUE)
library(ComplexHeatmap, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(WGCNA, quietly = TRUE)
library(tidyr, quietly = TRUE)
```

## - Loading palettes

In this section, we load additional color palettes and define some custom ones.

```{r message=FALSE, warning=FALSE}
# Load additional colour palette packages
library(unikn, quietly = TRUE)
library(RColorBrewer, quietly = TRUE)
library(yarrr, quietly = TRUE)
library(scales, quietly = TRUE)
library(ggsci, quietly = TRUE)

# Define a set of custom color palettes from the unikn package
mix_1 <- usecol(pal = c(Karpfenblau, "white", Peach), n = 15)
mix_2 <- usecol(pal = c(rev(pal_seeblau), "white", pal_pinky))
mix_3 <- usecol(pal = c(rev(pal_bordeaux), "white", pal_petrol), n = 15)

# Display the custom palettes
seecol(list(mix_1, mix_2, mix_3), col_brd = "white", lwd_brd = 4, title = "Comparing palettes mixed from unikn colors", pal_names = c("mix_1", "mix_2", "mix_3"))

# Define a second set of custom palettes from the RColorBrewer and yarrr packages
brew_mix <- usecol(c(rev(brewer.pal(n = 4, name = "Reds")), "white", brewer.pal(n = 4, name = "Blues")), n = 13)
brew_ext <- usecol(brewer.pal(n = 11, name = "Spectral"), n = 12)
yarrr_mix <- usecol(c(piratepal("nemo"), piratepal("bugs")))
yarrr_mod <- usecol(c(piratepal("ipod")), n = 9)

# Display the second set of custom palettes
seecol(pal = list(brew_mix, brew_ext, yarrr_mix, yarrr_mod), col_brd = "white", lwd_brd = 2, title = "Using usecol() and seecol() to mix and modify palettes", pal_names = c("brew_mix", "brew_ext", "yarrr_mix", "yarrr_mod"))

# Define additional custom palettes from the scales package
natjournals_palette <- pal_npg("nrc")(10)
```

## - Log Session Info

Finally, we log the session information for reproducibility.

```{r message=TRUE, warning=TRUE}
# Write the session information to a text file
writeLines(capture.output(sessionInfo()), file.path(script_folder, 'Fig_06_DEF_SessionInfo.txt'))

# Print the session information
sessionInfo()
```

# 2) Loading input files

## - Loading RF clusters

```{r}
RF_cluster_metadata <- read.csv(paste(data_folder, '/RandomForest/Results/Clustering/K2/RF_K2_Clustered_Metadata.csv', 
                                    sep = ''), row.names = 'Patient.ID')
```

## - Subsetting WGCNA modules enrichments

```{r}
WGCNA_score <- RF_cluster_metadata %>% select(ME1, ME2, ME3, ME4, ME5, ME6)
```

## - Subsetting common drugs resistance/response modules enrichments

```{r}
Common_drugs_response <- RF_cluster_metadata %>% select(AZA.resistance, Cytarabine.response, 
                                                        Daunorubicin.resistance, Venetoclax.resistance)

colnames(Common_drugs_response) <- c('Azacitidine resistance',
                                     'Cytarabine response',
                                     'Daunorubicin resistance',
                                     'Venetoclax resistance')
```

```{r}
Cor_df <- cbind(WGCNA_score, Common_drugs_response)
```

# 3) Cor. Genotype to drug response per subtype : Panel F

## - Data wrangling

This code creates model matrices for two different clusters of data, calculates correlations and p-values between these matrices and subsets of another data frame, transforms these p-values into a different scale, and reshapes the data for plotting. The transformed p-values are used to create a new variable for coloring a plot.

```{r}
# Create model matrices for clusters C1 and C2
model_matrix_C1 <- as.data.frame(model.matrix( ~ DNMT3A + TET2 + SRSF2 + SF3B1 + SETBP1 + U2AF1 + IDH1 + IDH2 + ASXL1 + EZH2 + STAG2 + BCOR + TP53 + PHF6 + NF1 + RUNX1 + FLT3 + NPM1 + CEBPA + WT1 + KIT + BRAF + KRAS + NRAS + CBL + PTPN11 + GATA2 + SH2B3 + ZRSR2 + CSF3R + DTA.mutations..DNMT3A..TET2..ASXL1. +Splicing.mutations..U2...SRSF2..SF3B1..U2AF1..ZRSF2. +RAS.mutations..NRAS..KRAS..CBL..NF1..PTPN11..BRAF. + IDH.mutations..IDH1..IDH2. - 1 , data = RF_cluster_metadata %>% filter(RF_Cluster == 'C1')))
model_matrix_C2 <- as.data.frame(model.matrix( ~ DNMT3A + TET2 + SRSF2 + SF3B1 + SETBP1 + U2AF1 + IDH1 + IDH2 + ASXL1 + EZH2 + STAG2 + BCOR + TP53 + PHF6 + NF1 + RUNX1 + FLT3 + NPM1 + CEBPA + WT1 + KIT + BRAF + KRAS + NRAS + CBL + PTPN11 + GATA2 + SH2B3 + ZRSR2 + CSF3R + DTA.mutations..DNMT3A..TET2..ASXL1. +Splicing.mutations..U2...SRSF2..SF3B1..U2AF1..ZRSF2. +RAS.mutations..NRAS..KRAS..CBL..NF1..PTPN11..BRAF. + IDH.mutations..IDH1..IDH2. - 1 , data = RF_cluster_metadata %>% filter(RF_Cluster == 'C2')))

# Rename the columns of the model matrices
colnames(model_matrix_C1) <- c(colnames(model_matrix_C1[,1:30]), 'DTA.mut.', 'Splicing.mut.', 'RAS.mut.', 'IDH.mut.')
colnames(model_matrix_C2) <- c(colnames(model_matrix_C1[,1:30]), 'DTA.mut.', 'Splicing.mut.', 'RAS.mut.', 'IDH.mut.')

# Calculate correlations and p-values for cluster C1
Cor_modules_C1 <- WGCNA::cor(model_matrix_C1, Cor_df[grep(TRUE, RF_cluster_metadata$RF_Cluster == 'C1'),7:10], use = "p")
Cor_modules_stat_C1 <- WGCNA::corPvalueStudent(Cor_modules_C1, nrow(model_matrix_C1))
Cor_modules_stat_C1_rank <- as.data.frame(sign(Cor_modules_C1)*(-1*log10(Cor_modules_stat_C1)))
Cor_modules_stat_C1_rank$Mutation <- rownames(Cor_modules_stat_C1_rank)
Cor_modules_stat_C1_rank_ggplot_df <- pivot_longer(as.data.frame(Cor_modules_stat_C1_rank), cols = c('Azacitidine resistance', `Cytarabine response`, `Daunorubicin resistance`, `Venetoclax resistance`))
Cor_modules_stat_C1_rank_ggplot_df$value_to_colour <- Cor_modules_stat_C1_rank_ggplot_df$value
Cor_modules_stat_C1_rank_ggplot_df[grep(TRUE,Cor_modules_stat_C1_rank_ggplot_df$value_to_colour < 1.30103 & Cor_modules_stat_C1_rank_ggplot_df$value_to_colour > -1.30103),'value_to_colour'] <- NA

# Calculate correlations and p-values for cluster C2
Cor_modules_C2 <- WGCNA::cor(model_matrix_C2, Cor_df[grep(TRUE, RF_cluster_metadata$RF_Cluster == 'C2'),7:10], use = "p")
Cor_modules_stat_C2 <- WGCNA::corPvalueStudent(Cor_modules_C2, nrow(model_matrix_C2))
Cor_modules_stat_C2_rank <- as.data.frame(sign(Cor_modules_C2)*(-1*log10(Cor_modules_stat_C2)))
Cor_modules_stat_C2_rank$Mutation <- rownames(Cor_modules_stat_C2_rank)
Cor_modules_stat_C2_rank_ggplot_df <- pivot_longer(as.data.frame(Cor_modules_stat_C2_rank), cols = c('Azacitidine resistance', `Cytarabine response`, `Daunorubicin resistance`, `Venetoclax resistance`))
Cor_modules_stat_C2_rank_ggplot_df$value_to_colour <- Cor_modules_stat_C2_rank_ggplot_df$value
Cor_modules_stat_C2_rank_ggplot_df[grep(TRUE,Cor_modules_stat_C2_rank_ggplot_df$value_to_colour < 1.30103 & Cor_modules_stat_C2_rank_ggplot_df$value_to_colour > -1.30103),'value_to_colour'] <- NA
```

## - Plotting

### - C1: Immature

This code creates a scatter plot of the transformed p-values against the names of the variables, with the points colored according to the transformed p-values. It also adds vertical lines to the plot, sets the theme and labels of the plot, adds text to the plot, and sets the color gradient of the plot. The plot is then printed and saved as a PNG file.

```{r echo=TRUE, fig.align='center', fig.cap='Figure 6F', fig.pos='H', fig.height=3.333, fig.width=6.254}
# Create a ggplot object
plot <- ggplot(data = Cor_modules_stat_C1_rank_ggplot_df, aes(x=value, y = name, colour = value_to_colour)) + 
  # Add points to the plot
  geom_point(size=3, stroke=0.2) + 
  # Add vertical lines to the plot
  geom_vline(xintercept = 1.30103, linetype="dashed", alpha = 0.45, color = "gray44", size=0.75) +  
  geom_vline(xintercept = -1.30103, linetype="dashed", alpha = 0.45, color = "gray44", size=0.75) +  
  # Set the theme of the plot
  theme_light(base_rect_size = 0) + 
  theme(legend.position = 'none', 
        panel.grid.major.y = element_line(color = "gray22", size = 0.35),
        panel.grid.major.x = element_line(color = "gray60", size = 0),
        panel.grid.minor.x = element_line(color = "gray60", size = 0),
        axis.text.y = element_text(size = 12, face= 'bold'),
        axis.title.x = element_text(size = 12, face= 'bold', hjust = 0.5),
        plot.title = element_text(size = 14, face= 'bold', hjust = 0.5),
        plot.subtitle = element_text(size = 12, hjust = 0.5)) +
  # Set the labels of the plot
  xlab(bquote(~Sign. -Log[10]~ '(P.Value)')) + ylab('') + 
  ggtitle(label = 'Genotype to drug response cor.', subtitle = 'C1: Immature') +
  # Add text to the plot
  geom_text_repel(aes(label = ifelse(value > 1.30103, as.character(Mutation), '')),
                  size = 4,
                  nudge_x = -0.05, 
                  nudge_y = 0.35, 
                  segment.size = 0.5,
                  direction = 'x') +
  geom_text_repel(aes(label = ifelse(value < -1.30103, as.character(Mutation), '')),
                    size = 4,
                    nudge_x = -0.15, 
                    nudge_y = -0.25, 
                    segment.size = 1,
                    direction = 'x') +
  # Set the color gradient of the plot
  scale_colour_gradient(low = '#2171B5', high = '#CB181D', na.value = ggplot2::alpha('grey72', 1)) +
  # Set the position of the y-axis
  scale_y_discrete(position = "right")

# Print the plot
print(plot)

# Save the plot as a PNG file
png(filename = paste(figures_folder, '/Fig_06/Panel_F-1.png', sep = ''), units = "px", width = 1876.2, height = 1000, res = 300)
print(plot)
invisible(dev.off())
```

### - C2: Mature

This code creates a scatter plot of the transformed p-values against the names of the variables, with the points colored according to the transformed p-values. It also adds vertical lines to the plot, sets the theme and labels of the plot, adds text to the plot, and sets the color gradient of the plot. The plot is then printed and saved as a PNG file.

```{r echo=TRUE, fig.align='center', fig.cap='Figure 6F', fig.pos='H', fig.height=3.333, fig.width=6.254}
# Create a ggplot object
plot <- ggplot(data = Cor_modules_stat_C2_rank_ggplot_df, aes(x=value, y = name, colour = value_to_colour)) + 
  # Add points to the plot
  geom_point(size=3, stroke=0.2) + 
  # Add vertical lines to the plot
  geom_vline(xintercept = 1.30103, linetype="dashed", alpha = 0.45, color = "gray44", size=0.75) +  
  geom_vline(xintercept = -1.30103, linetype="dashed", alpha = 0.45, color = "gray44", size=0.75) +  
  # Set the theme of the plot
  theme_light(base_rect_size = 0) + 
  theme(legend.position = 'none', 
        panel.grid.major.y = element_line(color = "gray22", size = 0.35),
        panel.grid.major.x = element_line(color = "gray60", size = 0),
        panel.grid.minor.x = element_line(color = "gray60", size = 0),
        axis.text.y = element_text(size = 12, face= 'bold'),
        axis.title.x = element_text(size = 12, face= 'bold', hjust = 0.5),
        plot.title = element_text(size = 14, face= 'bold', hjust = 0.5),
        plot.subtitle = element_text(size = 12, hjust = 0.5)) +
  # Set the labels of the plot
  xlab(bquote(~Sign. -Log[10]~ '(P.Value)')) + ylab('') + 
  ggtitle(label = 'Genotype to drug response cor.', subtitle = 'C2: Mature') +
  # Add text to the plot
  geom_text_repel(aes(label = ifelse(value > 1.30103, as.character(Mutation), '')),
                  size = 4,
                  nudge_x = -0.025, 
                  nudge_y = 0.15, 
                  segment.size = 0.5,
                  direction = 'x') +
  geom_text_repel(aes(label = ifelse(value < -1.30103, as.character(Mutation), '')),
                    size = 4,
                    nudge_x = -0, 
                    nudge_y = -0.15, 
                    segment.size = 1,
                    direction = 'x') +
  # Set the color gradient of the plot
  scale_colour_gradient(low = '#2171B5', high = '#CB181D', na.value = ggplot2::alpha('grey72', 1)) +
  # Set the position of the y-axis
  scale_y_discrete(position = "right")

# Print the plot
print(plot)

# Save the plot as a PNG file
png(filename = paste(figures_folder, '/Fig_06/Panel_F-2.png', sep = ''), units = "px", width = 1876.2, height = 1000, res = 300)
print(plot)
invisible(dev.off())
```

# 4) Cor. WGCNA to drug response

## - Heatmap : Panel D

This code calculates the correlation and p-values between the columns of a data frame, creates a heatmap of the correlation, and saves the heatmap as a PNG file. It also saves the correlation and p-values as CSV files. The heatmap is colored according to the correlation values, and cells are marked with "*" or "•" based on the p-values.

```{r echo=TRUE, fig.align='center', fig.cap='Figure 6D', fig.pos='H', fig.height=2.75591, fig.width=4.33071}
# Calculate the correlation between the columns of the data frame
Cor_modules <- WGCNA::cor(Cor_df[,7:10], Cor_df[,1:6], use = "p")
# Calculate the p-values of the correlation
Cor_modules_stat <- WGCNA::corPvalueStudent(Cor_modules, nrow(RF_cluster_metadata))

# Define the color gradient for the heatmap
breaks_val <- seq(from = -1, to = 1, length.out = 13)
col_heatmap_guide <- colorRamp2(breaks = breaks_val,colors = rev(brew_mix))

# Set the seed for reproducibility
set.seed(123)

# Create a heatmap
Heatmap_genes <- Heatmap(Cor_modules, name = 'Correlation score\n(pearson)',
                            col = col_heatmap_guide,
                            cell_fun = function(j, i, x, y, w, h, fill) {
                              if(Cor_modules_stat[i, j] < 0.001) {
                                  grid.text("*", x, y, vjust = 0.65, gp = gpar(fontsize = 18))
                              } else if(Cor_modules_stat[i, j] < 0.05) {
                                  grid.text("•", x, y, vjust = 0.2, gp = gpar(fontsize = 13))
                              }},
                            border = TRUE,
                            show_row_dend = TRUE, column_names_side = 'bottom',
                            use_raster = FALSE, 
                            rect_gp = gpar(col = "white", lwd = 0),
                            raster_by_magick = TRUE,
                            column_title_gp = gpar(fontsize = 0, angle=45),
                            column_names_gp = gpar(fontsize = 7, hjust=0),
                            row_title_gp = gpar(fontsize = 0),
                            row_names_gp = gpar(fontsize = 9, hjust=0, face='bold'),
                            show_heatmap_legend = T,
                            heatmap_legend_param = list(
                                      legend_direction = "horizontal", 
                                      legend_width = unit(3, "cm")), row_names_side = 'right')

# Draw the heatmap
draw(Heatmap_genes, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")

# Save the heatmap as a PNG file
png(filename = paste(figures_folder, '/Fig_06/Panel_D.png', sep = ''), units = "cm", width = 11, height = 7, res = 300)
draw(Heatmap_genes, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")
invisible(dev.off())

# Save the correlation and p-values as CSV files
write.csv(x = Cor_modules, 
          file = paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/Pathway_Correlations/Correlation_WGCNA_modules_to_Common_drug_response__COR.csv', sep = ''), quote = F)
write.csv(x = Cor_modules_stat, 
          file = paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/Pathway_Correlations/Correlation_WGCNA_modules_to_Common_drug_response__STATS.csv', sep = ''), quote = F)
```

## - Cor. WGCNA M5 to Venetoclax resistance : Panel E-1

This code loads two CSV files containing enrichment scores for WGCNA modules and Venetoclax resistance, respectively, against all samples including healthies.

```{r}
# Loading WGCNA Modules enrichment scores including healthy donors
WGCNA_score <- read.csv(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/WGCNA/WGCNA_modules__Enrichments_w_healthies.csv', 
                                    sep = ''), row.names = 'X')

# Loading Venetoclax resistance enrichment scores including healthy donors
Venetoclax_resistance <- read.csv(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Resistance/2020__Zhang_et_al__Ven_resistance_signature__Enrichments_w_healthies.csv', 
                                    sep = ''), row.names = 'X')

colnames(Venetoclax_resistance) <- 'Venetoclax resistance'
```

This code combines two data frames, adds a new column 'Cluster', and updates its values based on another data frame. It then creates a scatter plot of the 'ME5' column against the 'Venetoclax resistance' column, adds a grid, calculates and displays the correlation, and adds points to the plot with different colors based on the 'Cluster' column. The plot is then saved as a PNG file.

```{r echo=TRUE, fig.align='center', fig.cap='Figure 6E', fig.pos='H', fig.height=2.2800409449, fig.width=2.4477917323}
# Combine the WGCNA score and Venetoclax resistance data frames
Cor_df <- cbind(WGCNA_score, Venetoclax_resistance)
# Add a new column 'Cluster' and set its default value to 'Control'
Cor_df$Cluster <- 'Control'
# Update the 'Cluster' values for the rows that exist in the RF_cluster_metadata data frame
Cor_df[rownames(RF_cluster_metadata),'Cluster'] <- RF_cluster_metadata[,'RF_Cluster']

# Create a scatter plot with correlation
plot <- ggplot(data = Cor_df, mapping = aes(x = ME5, y = `Venetoclax resistance`)) + sm_hvgrid() + 
  sm_statCorr(color = 'black', corr_method = 'spearman') + 
  ylab('Venetoclax\nresistance') + xlab('WGCNA ME5') +
  theme_classic(base_line_size = 0.25) +
  theme(axis.title.x = element_text(face='bold', size = 11),
        axis.title.y = element_text(face='bold', size = 11),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9), 
        legend.position = 'none')

# Add points to the plot with different colors based on the 'Cluster' column
plot <- plot + geom_point(data = Cor_df, aes(fill=Cluster), 
                  shape = 21, color = 'white', size = 3) +
  scale_fill_manual(values = c('Control' = '#A4BECE',
                               'C1'    = '#793F0B',
                               'C2'    = '#0B544B'))

print(plot)

# Save the plot as a PNG file
png(filename = paste(figures_folder, '/Fig_06/Panel_E-1.png', sep = ''), units = "cm", width = 6.217391, height = 5.791304, res = 300)
print(plot)
invisible(dev.off())
```

## - Cor. WGCNA M3 to Cytarabine response : Panel E-2

This code loads two CSV files containing enrichment scores for WGCNA modules and Cytarabine response, respectively, against all samples including healthies.

```{r}
# Load the Cytarabine response enrichment scores including healthy donors
Cytarabine_response <- read.csv(paste(data_folder, '/RNASeq/InHouse/Pathway_Analysis/ssEnrichments/Drug_Response/2019__Xu_et_al_cytarabine_response_signature__Enrichments_w_healthies.csv', 
                                    sep = ''), row.names = 'X')

# Rename the column of the Cytarabine response data frame
colnames(Cytarabine_response) <- 'Cytarabine response'
```

This code combines two data frames, adds a new column 'Cluster', and updates its values based on another data frame. It then creates a scatter plot of the 'ME3' column against the 'Cytarabine response' column, adds a grid, calculates and displays the correlation, and adds points to the plot with different colors based on the 'Cluster' column. The plot is then saved as a PNG file.

```{r echo=TRUE, fig.align='center', fig.cap='Figure 6E', fig.pos='H', fig.height=2.2800409449, fig.width=2.4477917323}
# Combine the WGCNA score and Cytarabine response data frames
Cor_df <- cbind(WGCNA_score, Cytarabine_response)
# Add a new column 'Cluster' and set its default value to 'Control'
Cor_df$Cluster <- 'Control'
# Update the 'Cluster' values for the rows that exist in the RF_cluster_metadata data frame
Cor_df[rownames(RF_cluster_metadata),'Cluster'] <- RF_cluster_metadata[,'RF_Cluster']

# Create a scatter plot with correlation
plot <- ggplot(data = Cor_df, mapping = aes(x = ME3, y = `Cytarabine response`)) + sm_hvgrid() + 
  sm_statCorr(color = 'black', corr_method = 'spearman') + 
  ylab('Cytarabine\nresponse') + xlab('WGCNA ME3') +
  theme_classic(base_line_size = 0.25) +
  theme(axis.title.x = element_text(face='bold', size = 11),
        axis.title.y = element_text(face='bold', size = 11),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9), 
        legend.position = 'none')

# Add points to the plot with different colors based on the 'Cluster' column
plot <- plot + geom_point(data = Cor_df, aes(fill=Cluster), 
                  shape = 21, color = 'white', size = 3) +
  scale_fill_manual(values = c('Control' = '#A4BECE',
                               'C1'    = '#793F0B',
                               'C2'    = '#0B544B'))

print(plot)

# Save the plot as a PNG file
png(filename = paste(figures_folder, '/Fig_06/Panel_E-2.png', sep = ''), units = "cm", width = 6.217391, height = 5.791304, res = 300)
print(plot)
invisible(dev.off())
```
