---
title: "Figure 04 - Panels ACDEFG"
output: 
  rmdformats::robobook:
    lightbox: true
    fig_caption: true
---

```{=html}
<style type="text/css">
  body{
  font-size: 10pt;
}
</style>
```

# 1) Setup

## - Defining work directory

In this section, we define the working directory for the R Markdown document.

```{r setup, include=TRUE}
# Get the path of the current script
# Then get the parent directory of the parent directory of the parent directory
local_wd_folder <- dirname(dirname(dirname(rstudioapi::getSourceEditorContext()$path)))

# Set the root directory for knitr to the local working directory
knitr::opts_knit$set(root.dir = local_wd_folder)
```

## - Defining input data and output directories

Here, we define the directories for input data and output files.

```{r}
# Get the directory of the current script
script_folder <- dirname(rstudioapi::getSourceEditorContext()$path)

# Define the data folder and output folder
data_folder <- './Data'

dir.create(paste(data_folder, '/RandomForest/Results', sep = ''), showWarnings = FALSE)
dir.create(paste(data_folder, '/RandomForest/Results/Model', sep = ''), showWarnings = FALSE)
dir.create(paste(data_folder, '/RandomForest/Results/Clustering', sep = ''), showWarnings = FALSE)
dir.create(paste(data_folder, '/RandomForest/Results/Clustering/K2', sep = ''), showWarnings = FALSE)

#results_folder <- './Results'
figures_folder <- './Figures'
```

## - Setting seed

Setting a seed ensures that any random processes are reproducible.

```{r}
# Set a seed for reproducibility
set.seed(123)
```

## - Packages installation (optional)

In this section, we ensure that all necessary packages are installed.

```{r}
# Ensure BiocManager is available for installation of Bioconductor packages
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
BiocManager::install(version = "3.18", ask = FALSE)

# Define a list of required packages used in this script
packages_required <- c("coin", "randomForest", "dplyr", "circlize", "ComplexHeatmap", 
                       "dendextend", "ggrepel", "ggplot2", "rstatix", "tidyverse",
                       "unikn", "RColorBrewer", "yarrr", "scales", "ggsci")

# Identify any required packages that are not installed
packages_uninstalled <- packages_required[!(packages_required %in% installed.packages()[,"Package"])]

# Install any uninstalled packages
if(length(packages_uninstalled)) BiocManager::install(packages_uninstalled, ask = FALSE)
```

## - Loading packages

Here, we load the necessary packages for our analysis.

```{r message=FALSE, warning=FALSE}
library(coin, quietly = TRUE)
library(rstatix, quietly = TRUE)
library(randomForest, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(circlize, quietly = TRUE)
library(ComplexHeatmap, quietly = TRUE)
library(dendextend, quietly = TRUE)
library(ggrepel, quietly = TRUE)
library(ggplot2, quietly = TRUE)
```

## - Loading palettes

In this section, we load additional color palettes and define some custom ones.

```{r message=FALSE, warning=FALSE}
# Load additional colour palette packages
library(unikn, quietly = TRUE)
library(RColorBrewer, quietly = TRUE)
library(yarrr, quietly = TRUE)
library(scales, quietly = TRUE)
library(ggsci, quietly = TRUE)

# Define a set of custom color palettes from the unikn package
mix_1 <- usecol(pal = c(Karpfenblau, "white", Peach), n = 15)
mix_2 <- usecol(pal = c(rev(pal_seeblau), "white", pal_pinky))
mix_3 <- usecol(pal = c(rev(pal_bordeaux), "white", pal_petrol), n = 15)

# Display the custom palettes
seecol(list(mix_1, mix_2, mix_3), col_brd = "white", lwd_brd = 4, title = "Comparing palettes mixed from unikn colors", pal_names = c("mix_1", "mix_2", "mix_3"))

# Define a second set of custom palettes from the RColorBrewer and yarrr packages
brew_mix <- usecol(c(rev(brewer.pal(n = 4, name = "Reds")), "white", brewer.pal(n = 4, name = "Blues")), n = 13)
brew_ext <- usecol(brewer.pal(n = 11, name = "Spectral"), n = 12)
yarrr_mix <- usecol(c(piratepal("nemo"), piratepal("bugs")))
yarrr_mod <- usecol(c(piratepal("ipod")), n = 9)

# Display the second set of custom palettes
seecol(pal = list(brew_mix, brew_ext, yarrr_mix, yarrr_mod), col_brd = "white", lwd_brd = 2, title = "Using usecol() and seecol() to mix and modify palettes", pal_names = c("brew_mix", "brew_ext", "yarrr_mix", "yarrr_mod"))

# Define additional custom palettes from the scales package
natjournals_palette <- pal_npg("nrc")(10)
```

## - Log Session Info

Finally, we log the session information for reproducibility.

```{r message=TRUE, warning=TRUE}
# Write the session information to a text file
writeLines(capture.output(sessionInfo()), file.path(script_folder, 'Fig_04_ACDEFG_SessionInfo.txt'))

# Print the session information
sessionInfo()
```

# 2) Panel A

## - Loading input files

```{r}
Immunophenotype_data <- read.csv(paste(data_folder, '/Immunophenotype/CHR_BP_CMML_immunophenotype_common_markers.csv', sep = ''), row.names = 'Patient.ID')

# We discard PT426 for the following analysis as RNASeq did not work for this patient's sample
Immunophenotype_data <- Immunophenotype_data[setdiff(rownames(Immunophenotype_data), '426'),]
```

## - Heatmap plotting

Here we create a clustered heatmap of flow cytometry data, and save it as a PNG file.

```{r echo=TRUE, fig.align='center', fig.cap='Figure 4A', fig.pos='H', fig.height=4.33071, fig.width=7.87402}
# Convert the selected columns of the Immunophenotype_data dataframe to a matrix
flow_data_mat <- as.matrix(Immunophenotype_data[,c('SSCA','CD45','CD34','CD38',
                                                   'CD117','CD135','CD33','CD13',
                                                   'CD45RA','CD123','CD56')])

# Create a dataframe that classifies each column as either a 'Generic flow parameter' or a 'Cluster of differentiation'
identity_markers <- data.frame(row.names = colnames(flow_data_mat))
identity_markers[,'Class'] <- 'Cluster of differentiation'
identity_markers['SSCA','Class'] <- 'Generic flow parameter'
identity_markers$Class <- factor(identity_markers$Class, levels = c('Generic flow parameter', 'Cluster of differentiation'))

# Normalize the data to a range of 0 to 1
normalized = (flow_data_mat-min(flow_data_mat))/(max(flow_data_mat)-min(flow_data_mat))

# Scale the normalized data (mean of 0 and standard deviation of 1) and transpose it
scaled_flow_data_mat <- t(scale(as.matrix(normalized)))

# Define a sequence of break values for the heatmap color scale
breaks_val <- seq(from = -2, to = 2, length.out = 14)

# Create a color palette for the heatmap
col_heatmap_guide <- colorRamp2(breaks = breaks_val,colors = rev(c("#67000D",brew_mix)))

# Set a seed for reproducibility
set.seed(123)

# Create a row annotation for the heatmap, which will display a density plot for each row of data
row_anno = ComplexHeatmap::rowAnnotation(Distribution = anno_density(scaled_flow_data_mat, joyplot_scale = 1, 
                                                                     gp = gpar(fill = "#CCCCCC80")))

# Set the padding for the column and row annotations
ht_opt$COLUMN_ANNO_PADDING = unit(0.4, "cm")
ht_opt$ROW_ANNO_PADDING = unit(0.4, "cm")

# Create a heatmap
flow_heatmap <- Heatmap(scaled_flow_data_mat, 
                                        name = 'Scaled\nMFI', 
                                        col = col_heatmap_guide,
                                        border = T, 
                                        row_split = identity_markers$Class,
                                        cluster_columns = T, 
                                        cluster_rows = T, 
                                        row_title_gp = gpar(col='white', fontface = "bold"), 
                                        row_names_gp = gpar(fontface = "bold"), 
                                        rect_gp = gpar(col = "black", lwd = 0.05), 
                                        heatmap_legend_param = list(
                                          legend_direction = "horizontal", 
                                          legend_width = unit(5, "cm")), 
                                        row_names_side = 'left')

# Draw the heatmap
draw(flow_heatmap, heatmap_legend_side = "bottom", annotation_legend_side = "bottom")

# Save the heatmap as a PNG file
png(filename = paste(figures_folder, '/Fig_04/Panel_A.png', sep = ''), units = 'cm', width = 20, height = 11, res = 300)

# Draw the heatmap and its legends again, this time with the legends at the bottom of the plot
draw(flow_heatmap, heatmap_legend_side = "bottom", annotation_legend_side = "bottom")

# Shut down the graphics device
invisible(dev.off())
```

# - Panel CD

## - Loading input files

```{r}
RF.metadata <- read.csv(paste(data_folder, '/RandomForest/RF_input_data.csv', sep = ''))
```

## - Data wrangling

### - Assigning rownames

This code first converts the 'Patient.ID' column of the RF.metadata dataframe to a factor. It then sets the row names of the RF.metadata dataframe to the 'Patient.ID' column. Finally, it removes the 'Patient.ID' column from the RF.metadata dataframe.

```{r}
# Convert the 'Patient.ID' column of the RF.metadata dataframe to a factor
RF.metadata$Patient.ID <- as.factor(RF.metadata$Patient.ID)

# Set the row names of the RF.metadata dataframe to the 'Patient.ID' column
rownames(RF.metadata) <- RF.metadata$Patient.ID

# Remove the 'Patient.ID' column from the RF.metadata dataframe
RF.metadata$Patient.ID <- NULL
```

### - Converting data categories to factors or integers

Here we convert multiple columns in the RF.metadata dataframe to factors. Factors are used represent categorical data and can be used for statistical modeling.

```{r}
# Convert the specified columns of the RF.metadata dataframe to factors
RF.metadata$Gender <- as.factor(RF.metadata$Gender)
RF.metadata$Subtype.at.CMML <- as.factor(RF.metadata$Subtype.at.CMML)
RF.metadata$ICC.classification <- as.factor(RF.metadata$ICC.classification)
RF.metadata$Cutaneous.BPDCN <- as.factor(RF.metadata$Cutaneous.BPDCN)

RF.metadata$DNMT3A <- as.factor(RF.metadata$DNMT3A)
RF.metadata$TET2 <- as.factor(RF.metadata$TET2)
RF.metadata$SRSF2 <- as.factor(RF.metadata$SRSF2)
RF.metadata$SF3B1 <- as.factor(RF.metadata$SF3B1)
RF.metadata$SETBP1 <- as.factor(RF.metadata$SETBP1)
RF.metadata$U2AF1 <- as.factor(RF.metadata$U2AF1)
RF.metadata$IDH1 <- as.factor(RF.metadata$IDH1)
RF.metadata$IDH2 <- as.factor(RF.metadata$IDH2)
RF.metadata$ASXL1 <- as.factor(RF.metadata$ASXL1)
RF.metadata$EZH2 <- as.factor(RF.metadata$EZH2)
RF.metadata$STAG2 <- as.factor(RF.metadata$STAG2)
RF.metadata$BCOR <- as.factor(RF.metadata$BCOR)
RF.metadata$PHF6 <- as.factor(RF.metadata$PHF6)
RF.metadata$NF1 <- as.factor(RF.metadata$NF1)
RF.metadata$RUNX1 <- as.factor(RF.metadata$RUNX1)
RF.metadata$FLT3 <- as.factor(RF.metadata$FLT3)
RF.metadata$NPM1 <- as.factor(RF.metadata$NPM1)
RF.metadata$CEBPA <- as.factor(RF.metadata$CEBPA)
RF.metadata$WT1 <- as.factor(RF.metadata$WT1)
RF.metadata$KIT <- as.factor(RF.metadata$KIT)
RF.metadata$BRAF <- as.factor(RF.metadata$BRAF)
RF.metadata$KRAS <- as.factor(RF.metadata$KRAS)
RF.metadata$NRAS <- as.factor(RF.metadata$NRAS)
RF.metadata$CBL <- as.factor(RF.metadata$CBL)
RF.metadata$PTPN11 <- as.factor(RF.metadata$PTPN11)
RF.metadata$GATA2 <- as.factor(RF.metadata$GATA2)
RF.metadata$SH2B3 <- as.factor(RF.metadata$SH2B3)
RF.metadata$ZRSR2 <- as.factor(RF.metadata$ZRSR2)
RF.metadata$CSF3R <- as.factor(RF.metadata$CSF3R)
RF.metadata$DTA.mutations..DNMT3A..TET2..ASXL1. <- as.factor(RF.metadata$DTA.mutations..DNMT3A..TET2..ASXL1.)
RF.metadata$RAS.mutations..NRAS..KRAS..CBL..NF1..PTPN11..BRAF. <- as.factor(RF.metadata$RAS.mutations..NRAS..KRAS..CBL..NF1..PTPN11..BRAF.)
RF.metadata$IDH.mutations..IDH1..IDH2. <- as.factor(RF.metadata$IDH.mutations..IDH1..IDH2.)
RF.metadata$Splicing.mutations..U2...SRSF2..SF3B1..U2AF1..ZRSF2. <- as.factor(RF.metadata$Splicing.mutations..U2...SRSF2..SF3B1..U2AF1..ZRSF2.)

RF.metadata$Cytogenetics..at.transformation. <- as.factor(RF.metadata$Cytogenetics..at.transformation.)
```

## - Generating RandomForest model

This code first sets a seed for reproducibility. It then trains a random forest model on the RF.metadata dataframe, using 100,000 trees. The model is trained with proximity measures and variable importance measures. After training the model, the seed is reset. Finally, the trained model is saved to a .RDS file.

```{r}
# Set the seed for reproducibility
set.seed(123)

# Define the number of trees to be used in the random forest
nTree_iterations <- c(100000)

# Train a random forest model on the RF.metadata dataframe
# The proximity argument is set to TRUE to compute proximity measures for the observations
# The ntree argument specifies the number of trees to grow in the forest
# The importance argument is set to TRUE to compute variable importance measures
model <- randomForest(RF.metadata,
                      proximity = TRUE,
                      ntree = nTree_iterations, 
                      importance = TRUE)

# Reset the seed after running the random forest, as it changes the seed
set.seed(123)

# Save the trained model to a .RDS file
saveRDS(model, 
        file = paste(data_folder, '/RandomForest/Results/Model/RF_Classification_Model.RDS', sep = ''))
```

## - Analysis of RF model

### - Evaluating GINI Importance

This code first computes the Gini importance of each predictor in the model, sorts the importances in decreasing order, and converts the result to a data frame. It then adds a 'Feature_class' column to the data frame, indicating the type of each predictor. Features metrics were illustrated via GraphPad (see under ***GraphPad/Fig_04/Panels_CEFG.pzfx***).

```{r}
# Compute the Gini importance of each predictor in the model, sort them in decreasing order, and convert the result to a data frame
GINI <- suppressWarnings(data.frame(GINI_Index = sort(importance(model)[, "MeanDecreaseGini"], decreasing = TRUE)) %>% dplyr::add_rownames() %>% dplyr::rename(Predictor = 'rowname'))

# Convert GINI to a data frame
GINI <- as.data.frame(GINI)

# Set the row names of GINI to the 'Predictor' column
rownames(GINI) <- GINI$Predictor

# Add a 'Feature_class' column to GINI, indicating the type of each predictor
GINI[colnames(RF.metadata)[1:46],'Feature_class'] <- 'Clinical metadata'
GINI[colnames(RF.metadata)[47:57],'Feature_class'] <- 'Cytometry data'
GINI[colnames(RF.metadata)[58:177],'Feature_class'] <- 'Transcriptional data'

# Write the GINI data frame to a .csv file
write.table(x = GINI, 
            file = paste(data_folder, '/RandomForest/Results/Model/RF_Classification_Model_GINI_Importance.csv', sep = ''),
            quote = FALSE, row.names = F, col.names = TRUE, sep = ',')
```

### - Extracting proximity matrix for distance computation and MDS similarity evaluation

This script calculates a distance matrix from a Random Forest model's proximity matrix, performs multidimensional scaling (MDS) on the distance matrix, creates a data frame with the MDS values, and generates an MDS plot using ggplot2. The plot is labeled with the percentage of variance explained by each dimension.

```{r echo=TRUE, fig.align='center', fig.pos='H', fig.height=8, fig.width=5}
# Calculate distance matrix from Random Forest model's proximity matrix
distance.matrix <- dist(1-model$proximity)

# Perform multidimensional scaling (MDS) on the distance matrix
mds.stuff <- cmdscale(distance.matrix, eig = TRUE, x.ret = TRUE)

# Calculate the percentage of variance explained by each dimension
mds.var.per <- round(mds.stuff$eig/sum(mds.stuff$eig)*100, 1)

# Extract the MDS values
mds.values <- mds.stuff$points

# Create a data frame with the MDS values
mds.data <- data.frame(Sample=rownames(mds.values),
                       X= mds.values[,1],
                       Y= mds.values[,2])

# Generate an MDS plot using ggplot2
ggplot(data = mds.data, aes(x = X, y = Y, label=Sample)) + 
  geom_text(aes(color=Sample)) +
  theme_bw() +
  xlab(paste("MDS1 - ", mds.var.per[1], "%", sep="")) +
  ylab(paste("MDS2 - ", mds.var.per[2], "%", sep="")) +
  ggtitle(label = "MDS plot", subtitle = "(1 - Random Forest Proximities)") + 
  theme(legend.position = 'bottom')
```

### - Clustering K = 2

#### - Hierarchical clustering on distance matrix : Panel D

This script generates a hierarchical clustering dendrogram from a distance matrix, saves it as a PNG file, cuts the dendrogram into two clusters, annotates the clusters, and creates a scatter plot of the MDS values colored by cluster. It then saves the scatter plot as a PNG file, adds the cluster annotations to the metadata, and writes the updated metadata to a CSV file.

```{r echo=TRUE, fig.align='center', fig.cap='Figure 4D', fig.pos='H', fig.height=4.133858, fig.width=8.26772}
# Generate hierarchical clustering dendrogram
dendro <- stats::hclust(distance.matrix, method = "ward.D2") %>% as.dendrogram %>% ladderize  %>% color_branches(k=2, col = c('#793F0B', '#0B544B')) %>% color_labels(k = 2, col = c('#793F0B', '#0B544B'))

# Plot dendrogram
par(lwd=3)
plot(dendro, axes=F)

# Save PNG
png(filename = paste(figures_folder, '/Fig_04/Panel_D-1.png', sep = ''), units = "cm", width = 21, height = 10.5, res = 300)

# Plot dendrogram
par(lwd=3)
plot(dendro, axes=F)

# Close PNG device
invisible(dev.off())
```

```{r echo=TRUE, fig.align='center', fig.cap='Figure 4D', fig.pos='H', fig.height=5.51181, fig.width=5.51181}
# Cut dendrogram into two clusters
Hierchical_clusters_K2 <- sort(cutree(dendro, k=2))

# Annotate clusters
Hierarchical_clustering_K2 <- data.frame(row.names = rownames(RF.metadata))
Hierarchical_clustering_K2$Patient.ID <- rownames(RF.metadata)
for (ID in Hierarchical_clustering_K2$Patient.ID) {
  Hierarchical_clustering_K2[ID,'Cluster'] <- Hierchical_clusters_K2[[ID]]
}
Hierarchical_clustering_K2[Hierarchical_clustering_K2$Cluster == Hierarchical_clustering_K2['416', 'Cluster'],'Cluster_annotated'] <- 'C1'
Hierarchical_clustering_K2[Hierarchical_clustering_K2$Cluster == Hierarchical_clustering_K2['55', 'Cluster'],'Cluster_annotated'] <- 'C2'

# Create scatter plot of MDS values colored by cluster
mds.data.Hierarchical_K2_clustered <- mds.data
mds.data.Hierarchical_K2_clustered$Cluster <- Hierarchical_clustering_K2$Cluster_annotated
cols <- c('C1' = '#793F0B', 'C2' = '#0B544B')

plot.Hierarchical_clusters_K2 <- ggplot(data = mds.data.Hierarchical_K2_clustered, aes(x = X, y = Y, label=Sample)) + 
  geom_point(aes(color=Cluster, fill=Cluster), shape = 21, alpha = 0.8, size = 2.5) +
  theme_linedraw(base_line_size = 0.2) +
  xlab("MDS1") +
  ylab("MDS2") +
  ggtitle(label = "") +
  scale_colour_manual(
    values = cols,
    aesthetics = c("colour", "fill")
  ) + theme(legend.position = 'none',
            axis.title.x = element_text(family='Helvetica', size=12, face = 'bold', vjust = 0),
            axis.title.y = element_text(family='Helvetica', size=12, face = 'bold', vjust = +3))

plot.Hierarchical_clusters_K2_labelled <- plot.Hierarchical_clusters_K2 + 
  geom_label_repel(aes(label = Sample, color=Cluster), 
                   box.padding = 0.65, point.padding = 0.1, 
                   size = 2, segment.color = 'grey70', max.overlaps = 15)

# Plotting the above chart
print(plot.Hierarchical_clusters_K2_labelled)

# Save scatter plot as PNG
png(filename = paste(figures_folder, '/Fig_04/Panel_D-2.png', sep = ''), units = "cm", width = 14, height = 14, res = 300)
print(plot.Hierarchical_clusters_K2_labelled)
invisible(dev.off())

# Add cluster annotations to metadata
RF.metadata.hierarchical_clustering_K2 <- RF.metadata
RF.metadata.hierarchical_clustering_K2[Hierarchical_clustering_K2$Patient.ID,'RF_Cluster'] <- Hierarchical_clustering_K2$Cluster_annotated

# Add the Patient.ID column back into the data frame
RF.metadata.hierarchical_clustering_K2 <- RF.metadata.hierarchical_clustering_K2 %>% mutate(.before = 'Gender', 'Patient.ID' = rownames(RF.metadata.hierarchical_clustering_K2))

# Write updated metadata to CSV
write_csv(x = RF.metadata.hierarchical_clustering_K2, 
            file = paste(data_folder, '/RandomForest/Results/Clustering/K2/RF_K2_Clustered_Metadata.csv', sep = ''))
```

#### - Wilcox t-test unpaired - median comparisons : Panels CEFG

This script converts all variables in the metadata to numeric, splits the metadata into two groups based on the cluster annotations, performs a Wilcoxon test for each variable to compare the two groups, and writes the test results to a CSV file. Features were then illustrated via GraphPad (see under ***GraphPad/Fig_04/Panels_CEFG.pzfx***).

```{r}
# Remove the Patient.ID column
RF.metadata.hierarchical_clustering_K2$Patient.ID <- NULL

# Convert all variables to numeric
RF.metadata.hierarchical_clustering_K2_numerised <- RF.metadata.hierarchical_clustering_K2[1:length(colnames(RF.metadata.hierarchical_clustering_K2))]
for (variable in colnames(RF.metadata.hierarchical_clustering_K2_numerised)[1:length(colnames(RF.metadata.hierarchical_clustering_K2))-1]) {
  RF.metadata.hierarchical_clustering_K2_numerised[,variable] <- as.numeric(RF.metadata.hierarchical_clustering_K2_numerised[,variable])
}

# Split metadata into two groups based on cluster annotations
Group_A <- RF.metadata.hierarchical_clustering_K2_numerised %>% dplyr::filter(RF_Cluster == 'C1')
Group_B <- RF.metadata.hierarchical_clustering_K2_numerised %>% dplyr::filter(RF_Cluster == 'C2')

# Perform Wilcoxon test for each variable
wilcox_t_test_results <- data.frame(row.names = colnames(RF.metadata.hierarchical_clustering_K2_numerised)[1:length(colnames(RF.metadata.hierarchical_clustering_K2))-1])
for (variable in colnames(RF.metadata.hierarchical_clustering_K2_numerised)[1:length(colnames(RF.metadata.hierarchical_clustering_K2))-1]) {
  p <- coin::wilcox_test(formula = get(variable) ~ as.factor(RF_Cluster), data=RF.metadata.hierarchical_clustering_K2_numerised, distribution = "exact")
  wilcox_t_test_results[variable,'pval'] <- coin::pvalue(p)
  wilcox_t_test_results[variable,'Median C1'] <- median(Group_A[,variable])
  wilcox_t_test_results[variable,'Median C2'] <- median(Group_B[,variable])
  wilcox_t_test_results[variable, paste('Min', 'C1', sep = ' ')] <- min(Group_A[,variable])
  wilcox_t_test_results[variable, paste('Min', 'C2', sep = ' ')] <- min(Group_B[,variable])
  wilcox_t_test_results[variable, paste('Max', 'C1', sep = ' ')] <- max(Group_A[,variable])
  wilcox_t_test_results[variable, paste('Max', 'C2', sep = ' ')] <- max(Group_B[,variable])
  wilcox_t_test_results[variable, paste('Std', 'C1', sep = ' ')] <- sd(Group_A[,variable])
  wilcox_t_test_results[variable, paste('Std', 'C2', sep = ' ')] <- sd(Group_B[,variable])
}

# Write test results to CSV
write.csv(wilcox_t_test_results, 
          file = paste(data_folder, '/RandomForest/Results/Clustering/K2/RF_K2_Stats_Wilcox_t_test.csv', sep = ''),
          row.names = T, quote = F)
```

